\textbf{APPENDIX}

\textbf{Glossary of the Seafood Sustainability Disclosure Ecosystem}

Frameworks, regulations, certifications and standards --- Layer 1
(demand) and Layer 2 (enablement).

\emph{Snapshot: June 2026. Regulatory references and statuses current as
of this date.}

Contents

How to use this glossary

This glossary catalogues every instrument in the seafood sustainability
landscape mapped in this thesis. It works both as an orientation read: what each actor in the ecosystem actually is and who runs it, and
as a reference appendix to the coverage matrix. Entries are grouped into
two layers, mirroring the demand analysis in the methodology.

\textbf{Layer 1 --- Demand-side instruments (disclosure askers).}
Instruments that place a demand directly on the firm: they require it to
report, certify, disclose or comply. These are the columns of the
cross-framework demand matrix, sub-grouped by the stakeholder that
drives them (EU horizontal, EU sector/operational, customer, investor,
industry/NGO).

\textbf{Layer 2 --- Enablement / methodology layer.} Instruments that
provide the methods, accounting rules, benchmarks or data standards that
make disclosure possible but do not themselves demand reporting from the
firm. They are documented in the landscape yet held separate from the
demand count --- central to the capability-corridor argument rather than
to the hotspot tally.

\textbf{Dual-role instruments.} A few sit on the boundary --- TCFD/TNFD
are frameworks that also shape investor asks; GDST is a data standard
buyers may require; the EU Taxonomy informs financiers. Each is placed
by its primary role, with the dual nature noted in its entry.

Each entry records, in a fixed metadata block: \textbf{the organisation
behind the instrument}, its legal reference or founding year, its
status, its stakeholder source and whether it is an asker or enabler,
followed by its objective, its role in the seafood ecosystem, and ---
where relevant --- an evidence note and an official source link.

\textbf{Reading the flags.} Following the project\textquotesingle s
evidence convention:  A
\textbf{Status flag} highlights instruments whose scope, timing or
legal standing was actively changing as of June 2026 --- chiefly the EU
\textquotesingle Omnibus\textquotesingle{} simplification (CSRD, CSDDD,
ESRS), the postponed EUDR timetable, the voluntary status of the VSME,
and the future application dates of the Forced Labour Regulation and
EmpCo.

Layer 1 --- Demand-side instruments (disclosure askers)

1A · EU horizontal regulation --- disclosure, due diligence \& finance

CSRD --- Corporate Sustainability Reporting Directive

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & EU (Commission proposal; co-legislated by
Parliament \& Council) \\
\textbf{Reference / founded} & Directive (EU) 2022/2464 --- in force
January 2023 \\
\textbf{Status · role} & EU directive; mandatory, transposed nationally
· Asker \\
\textbf{Stakeholder source} & Regulator (horizontal) \\
\textbf{Official source} &
\href{https://eur-lex.europa.eu/eli/dir/2022/2464/oj}{eur-lex.europa.eu
--- Directive (EU) 2022/2464} \\
\end{longtable}
}

\textbf{Objective.} Expands and standardises mandatory corporate
sustainability reporting across the EU, requiring in-scope undertakings
to report under the ESRS on a double-materiality basis (impacts on
people and planet, and sustainability matters financially material to
the firm).

\textbf{In the ecosystem.} Sets the scope boundary that decides whether
a seafood processor must report at all. Most 50--250-employee processors
fall below its thresholds and are reached only as value-chain partners
of larger in-scope customers.

\textbf{Status flag.} \emph{Scope and timing are being narrowed by the
2025 \textquotesingle Omnibus I\textquotesingle{} simplification package
(raising thresholds and delaying waves). Treat the in-scope population
as a moving target and verify the consolidated text before citing
thresholds.}

ESRS --- European Sustainability Reporting Standards

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & Developed by EFRAG (technical adviser); adopted by
the European Commission \\
\textbf{Reference / founded} & Commission Delegated Regulation (EU)
2023/2772 --- first set, 2023 \\
\textbf{Status · role} & Delegated regulation under CSRD; mandatory for
in-scope firms · Asker \\
\textbf{Stakeholder source} & Regulator (horizontal) \\
\textbf{Official source} &
\href{https://eur-lex.europa.eu/eli/reg_del/2023/2772/oj}{eur-lex.europa.eu
--- Del. Reg (EU) 2023/2772} \\
\end{longtable}
}

\textbf{Objective.} The detailed disclosure standards that
operationalise the CSRD: two cross-cutting standards (ESRS 1--2) plus
ten topical standards (E1--E5, S1--S4, G1), all built on double
materiality and a defined set of Disclosure Requirements (DRs).

\textbf{In the ecosystem.} The DR-level ruler against which this thesis
maps all cohort and SME disclosure; it defines the topical Disclosure
Requirements that the coverage matrix scores frameworks against.

\textbf{Status flag.} \emph{A simplified ESRS set (fewer datapoints)
is under EFRAG revision following the Omnibus mandate --- the exact
DR/datapoint counts may change.}

VSME --- Voluntary Sustainability Reporting Standard for SMEs

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & EFRAG (technical adviser to the European
Commission) \\
\textbf{Reference / founded} & Delivered to the Commission Dec 2024;
Commission Recommendation adopted 30 July 2025 \\
\textbf{Status · role} & Voluntary EU standard; non-binding (not a
delegated act, so no legal standing under CSRD Art. 29c yet) · Asker \\
\textbf{Stakeholder source} & Customer --- recoded from issuer: EFRAG
issues it, but its purpose is to let an SME answer its retail customers'
ESRS-cascade data requests with one standard; the value-chain channel
dominates, with lenders a secondary audience \\
\textbf{Official source} &
\href{https://ec.europa.eu/finance/docs/law/250730-recommendation-vsme_en.pdf}{ec.europa.eu
--- VSME Recommendation (30 July 2025)} \\
\end{longtable}
}

\textbf{Objective.} A proportionate voluntary standard for SMEs outside
CSRD scope, structured as a Basic Module and a Comprehensive Module,
designed to cut the reporting load from roughly a thousand ESRS
datapoints to around a hundred and to answer the most common value-chain
and lender information requests.

\textbf{In the ecosystem.} The realistic regulatory reference point ---
and candidate \textquotesingle home base\textquotesingle{} --- for
non-CSRD seafood SMEs; built specifically so a small supplier can
respond to a large customer\textquotesingle s ESRS-driven data request
once rather than per buyer.

\textbf{Status flag.} \emph{It is a Commission Recommendation, not a
Delegated Act --- adoption is encouraged but voluntary as of June 2026.
Do not describe it as legally binding.}

\textbf{Coding note.} VSME is the one instrument scored as two Phase~3
columns --- \emph{Basic} and \emph{Comprehensive} --- because its two
officially-defined modules impose different datapoint sets and the Basic
module is itself a complete, separately-usable SME report. The split is
warranted only where an instrument has officially-defined, nested
compliance tiers each forming a complete report (which is why GRASP,
SMETA's two-pillar variant and CDP's thematic modules are \emph{not}
split). Because the Comprehensive module contains the Basic module, the
two columns are nested rather than independent: the Comprehensive column
is coded on its \emph{incremental} datapoints only, and the pair carries a
single combined independence weight in the convergence metric so that VSME
is not double-counted.

CSDDD --- Corporate Sustainability Due Diligence Directive

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & EU \\
\textbf{Reference / founded} & Directive (EU) 2024/1760 \\
\textbf{Status · role} & EU directive; phased transposition · Asker \\
\textbf{Stakeholder source} & Customer (value chain) --- recoded from
Regulator: the Directive binds large customers, not the sub-threshold
processor, who experiences it as a cascaded contractual requirement \\
\textbf{Official source} &
\href{https://eur-lex.europa.eu/eli/dir/2024/1760/oj}{eur-lex.europa.eu
--- Directive (EU) 2024/1760} \\
\end{longtable}
}

\textbf{Objective.} Requires large companies to conduct human-rights and
environmental due diligence across their chains of activities ---
identifying, preventing, mitigating and accounting for adverse impacts.
It is a conduct/due-diligence directive rather than a reporting
standard.

\textbf{In the ecosystem.} Drives value-chain due-diligence expectations
(e.g. forced labour in fishing fleets) that cascade contractually onto
SME suppliers even when those SMEs are not themselves in scope.

\textbf{Status flag.} \emph{Materially amended by the Omnibus~I
Directive (published 26 Feb 2026, in force 18 Mar 2026): direct scope
narrowed to companies with $>$EUR 1.5\,bn turnover and 5{,}000
employees (\textasciitilde13{,}000 $\to$ \textasciitilde6{,}000 firms),
the climate-transition-plan obligation deleted, the harmonised civil-liability
regime removed, and due diligence recalibrated to a scoping exercise.
CSDDD-related transposition is due 26 Jul 2028 with application from 26
Jul 2029. The narrower direct scope reinforces, rather than weakens, the
indirect cascade locus used here: the SME is even less likely to be in
scope itself and is reached only through in-scope customers' contractual
requirements. The topic-imposition scoring is unaffected.}

EU Taxonomy --- Regulation on the establishment of a framework to
facilitate sustainable investment

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & EU \\
\textbf{Reference / founded} & Regulation (EU) 2020/852 \\
\textbf{Status · role} & EU regulation; directly applicable · Asker /
Enabler \\
\textbf{Stakeholder source} & Regulator (capital-market facing) \\
\textbf{Official source} &
\href{https://eur-lex.europa.eu/eli/reg/2020/852/oj}{eur-lex.europa.eu
--- Regulation (EU) 2020/852} \\
\end{longtable}
}

\textbf{Objective.} A classification system defining environmentally
sustainable economic activities against six environmental objectives,
using technical screening criteria,
\textquotesingle do-no-significant-harm\textquotesingle{} tests and
minimum social safeguards.

\textbf{In the ecosystem.} Shapes how banks and investors assess
\textquotesingle green\textquotesingle{} exposure; reaches seafood SMEs
mainly through lender and investor alignment questions rather than as a
direct obligation.

\textbf{Excluded from the scored matrix.} The Taxonomy is retained as
context but is not scored as a core coverage column. The reason is
specific to its mechanism, not to it being a finance instrument: the
Taxonomy operates through an \emph{activity-eligibility gate}, and the
focal activity --- seafood processing, NACE C10.20 --- is not a
Taxonomy-eligible activity in any current delegated act, including the
2023 Environmental Delegated Act. The gate therefore returns null, so the
Taxonomy generates no topic demand on the processor's core impacts (GHG,
water, circularity, biodiversity) and would only inflate coverage if
scored. This distinguishes it from the finance instruments that are
retained and scored --- SFDR (whose principal-adverse-impact datapoints
cascade to investees regardless of activity) and IFRS~S1/S2 (a
defined-datapoint disclosure standard) --- which impose topic demand
independently of activity eligibility. The Taxonomy re-enters only as a
\emph{project-conditional finance lens}: where the SME finances a specific
eligible CapEx/OpEx project (e.g.\ solar PV, heat pumps, building
renovation, EV charging, water infrastructure) or a lender requests
Taxonomy information, the eligibility, technical-screening,
do-no-significant-harm, minimum-safeguards and CapEx/OpEx-allocation
analysis attaches to the project, not to the seafood-processing business
as a whole.

SFDR --- Sustainable Finance Disclosure Regulation

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & EU \\
\textbf{Reference / founded} & Regulation (EU) 2019/2088 --- applies
from 2021 \\
\textbf{Status · role} & EU regulation; targets financial market
participants · Asker \\
\textbf{Stakeholder source} & Regulator (investor-facing) \\
\textbf{Official source} &
\href{https://eur-lex.europa.eu/eli/reg/2019/2088/oj}{eur-lex.europa.eu
--- Regulation (EU) 2019/2088} \\
\end{longtable}
}

\textbf{Objective.} Requires financial market participants and advisers
to disclose how they integrate sustainability risks and to report
principal adverse impacts (PAI) of their investments.

\textbf{In the ecosystem.} Indirect pressure on the firm: SFDR-bound
funds and banks pass PAI data requests down to investee and borrower
seafood companies.

1B · EU sector \& operational regulation --- seafood, food \& labour

CFP --- Common Fisheries Policy

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & EU \\
\textbf{Reference / founded} & Regulation (EU) No 1380/2013 \\
\textbf{Status · role} & EU regulation; directly applicable · Asker \\
\textbf{Stakeholder source} & Regulator (sector) \\
\textbf{Official source} &
\href{https://eur-lex.europa.eu/eli/reg/2013/1380/oj}{eur-lex.europa.eu
--- Reg (EU) 1380/2013} \\
\end{longtable}
}

\textbf{Objective.} The overarching EU framework for managing fishing
fleets and conserving stocks --- fishing-opportunity setting
(TACs/quotas), fleet-capacity ceilings, the landing obligation and a
maximum-sustainable-yield (MSY) exploitation objective.

\textbf{In the ecosystem.} Tier-1 legal licence to operate for
wild-capture sourcing; it binds Member States, and processor-facing
obligations flow downstream through the Control Regulation, IUU and CMO
rather than from the CFP itself.

IUU Regulation --- Illegal, Unreported and Unregulated Fishing
Regulation

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & EU \\
\textbf{Reference / founded} & Council Regulation (EC) No 1005/2008 ---
applies from 2010 \\
\textbf{Status · role} & EU regulation; directly applicable · Asker \\
\textbf{Stakeholder source} & Regulator (sector) \\
\textbf{Official source} &
\href{https://eur-lex.europa.eu/eli/reg/2008/1005/oj}{eur-lex.europa.eu
--- Reg (EC) 1005/2008} \\
\end{longtable}
}

\textbf{Objective.} Prevents, deters and eliminates illegal, unreported
and unregulated fishing, chiefly through a catch-certificate scheme:
imported wild-capture products must carry a catch certificate validated
by the flag state.

\textbf{In the ecosystem.} Anchor of the external traceability corridor
--- a validated catch certificate is a Tier-1 precondition for placing
imported wild fish on the EU market, complementary to (not a subset of)
the internal Control Regulation.

\textbf{Status flag.} \emph{The catch-certificate scheme is being
digitalised through the EU CATCH IT system, introduced by the 2023
control reform (Reg (EU) 2023/2842); roll-out is phased --- verify the
current application dates before citing.}

Control Regulation (recast) --- Fisheries Control Regulation

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & EU \\
\textbf{Reference / founded} & Regulation (EU) 2023/2842, amending Reg
(EC) 1224/2009 \\
\textbf{Status · role} & EU regulation; directly applicable (phased) ·
Asker \\
\textbf{Stakeholder source} & Regulator (sector) \\
\textbf{Official source} &
\href{https://eur-lex.europa.eu/eli/reg/2023/2842/oj}{eur-lex.europa.eu
--- Reg (EU) 2023/2842} \\
\end{longtable}
}

\textbf{Objective.} Modernises EU fisheries control and enforcement ---
including full-chain digital traceability, mandatory electronic catch
documentation and tighter monitoring of the EU fleet and the EU market.

\textbf{In the ecosystem.} Governs internal EU-fleet control and
EU-market traceability (the sibling to IUU\textquotesingle s external
import scheme); pushes lot-level digital traceability that feeds the
traceability-capability corridor.

\textbf{Status flag.} \emph{The recast applies on a phased timeline:
several provisions --- including the digitalised catch-documentation and
lot-level traceability obligations most relevant to processors --- enter
application on deferred dates. Verify the exact application dates before
citing them.}

CMO Regulation --- Common Market Organisation for fishery \& aquaculture
products

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & EU \\
\textbf{Reference / founded} & Regulation (EU) No 1379/2013 \\
\textbf{Status · role} & EU regulation; directly applicable · Asker \\
\textbf{Stakeholder source} & Regulator (sector) \\
\textbf{Official source} &
\href{https://eur-lex.europa.eu/eli/reg/2013/1379/oj}{eur-lex.europa.eu
--- Reg (EU) 1379/2013} \\
\end{longtable}
}

\textbf{Objective.} Organises the EU market for fishery and aquaculture
products, including common marketing standards and mandatory consumer
information --- commercial and scientific species name, production
method, catch/production area and gear category.

\textbf{In the ecosystem.} Binds the processor at the point of sale; its
mandatory origin and method labelling overlap directly with
chain-of-custody and consumer-information disclosures.

Fisheries Labour Directive --- Implementation of the ILO Work in Fishing
Convention (C188)

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & EU (implementing social-partner agreement on ILO
C188) \\
\textbf{Reference / founded} & Council Directive (EU) 2017/159 \\
\textbf{Status · role} & EU directive · Asker \\
\textbf{Stakeholder source} & Customer (value chain) --- recoded from
Regulator: the Directive binds vessel owners, not the processor, and
reaches the firm as value-chain content via buyer social audits
(e.g.\ SMETA), value-chain due diligence and ESRS~S2 \\
\textbf{Official source} &
\href{https://eur-lex.europa.eu/eli/dir/2017/159/oj}{eur-lex.europa.eu
--- Dir (EU) 2017/159} \\
\end{longtable}
}

\textbf{Objective.} Sets minimum working and living conditions for
fishers aboard vessels (contracts, hours/rest, repatriation,
accommodation, medical care).

\textbf{In the ecosystem.} Binds vessel owners rather than the
processor; reaches the firm as value-chain content via buyer SMETA
audits and ESRS S2, and as a baseline for forced-labour scrutiny.

\textbf{Status flag.} \emph{Considered for the scored set but EXCLUDED:
it is the only EU instrument linking fisheries labour directly, and
admitting it on value-chain-transmission grounds would oblige the
inclusion of the wider body of sector-specific labour law. Its
value-chain-labour signal is already carried by SMETA, the Forced Labour
Regulation and ESRS~S2. Retained in this glossary for completeness;
recorded as excluded in the scope log.}

EUDR --- EU Deforestation Regulation

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & EU \\
\textbf{Reference / founded} & Regulation (EU) 2023/1115 --- phased
application \\
\textbf{Status · role} & EU regulation; directly applicable · Asker \\
\textbf{Stakeholder source} & Regulator (sector-adjacent) \\
\textbf{Official source} &
\href{https://eur-lex.europa.eu/eli/reg/2023/1115/oj}{eur-lex.europa.eu
--- Reg (EU) 2023/1115} \\
\end{longtable}
}

\textbf{Objective.} Prohibits placing on (or exporting from) the EU
market certain commodities linked to deforestation --- cattle, soy, palm
oil, wood, cocoa, coffee, rubber --- without due-diligence statements
backed by geolocation evidence.

\textbf{In the ecosystem.} Relevant to aquaculture feed sourcing (soy
and palm derivatives) and to paper/board packaging, and to
natural-ecosystem-conversion topics.

\textbf{Status flag.} \emph{Application was postponed by 12 months:
large/medium operators from 30 Dec 2025 and micro/small operators from
30 June 2026 --- confirm the current date for your size class.}

PPWR --- Packaging and Packaging Waste Regulation

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & EU \\
\textbf{Reference / founded} & Regulation (EU) 2025/40 --- in force 11
Feb 2025 \\
\textbf{Status · role} & EU regulation; main obligations apply from 12
Aug 2026 · Asker \\
\textbf{Stakeholder source} & Regulator (horizontal) \\
\textbf{Official source} &
\href{https://eur-lex.europa.eu/eli/reg/2025/40/oj}{eur-lex.europa.eu
--- Reg (EU) 2025/40} \\
\end{longtable}
}

\textbf{Objective.} Replaces the 1994 Packaging Directive with
directly-applicable rules on packaging minimisation, recyclability,
recycled-content targets and labelling for all packaging placed on the
EU market.

\textbf{In the ecosystem.} Touches resource-use and waste (E5) plus the
packaging data that retailers increasingly request; binds the processor
as the entity placing packaged product on the market.

EmpCo --- Empowering Consumers for the Green Transition Directive

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & EU \\
\textbf{Reference / founded} & Directive (EU) 2024/825 --- in force 26
Mar 2024 \\
\textbf{Status · role} & EU directive; transpose by 27 Mar 2026, apply
from 27 Sep 2026 · Asker \\
\textbf{Stakeholder source} & Regulator (horizontal) \\
\textbf{Official source} &
\href{https://eur-lex.europa.eu/eli/dir/2024/825/oj}{eur-lex.europa.eu
--- Dir (EU) 2024/825} \\
\end{longtable}
}

\textbf{Objective.} Amends the Unfair Commercial Practices and Consumer
Rights Directives to ban generic, unsubstantiated green claims
(\textquotesingle eco-friendly\textquotesingle, \textquotesingle climate
neutral\textquotesingle) and to regulate sustainability labels.

\textbf{In the ecosystem.} Constrains
\textquotesingle sustainable\textquotesingle{} and eco claims on seafood
packaging and raises the evidentiary bar for certification-based
marketing.

\textbf{Status flag.} \emph{It is the consumer-law half of the EU
greenwashing agenda and the \emph{live} anti-greenwashing instrument; the
companion Green Claims Directive (substantiation rules) had its
withdrawal announced by the Commission in June 2025 (the final trilogue
was cancelled) and its status remained unresolved as of mid-2026 --- do
not conflate the two.}

Forced Labour Regulation --- Prohibition of products made with forced
labour on the Union market

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & EU \\
\textbf{Reference / founded} & Regulation (EU) 2024/3015 --- published
Dec 2024 \\
\textbf{Status · role} & EU regulation; applies from 14 Dec 2027 ·
Asker \\
\textbf{Stakeholder source} & Regulator (horizontal) \\
\textbf{Official source} &
\href{https://eur-lex.europa.eu/eli/reg/2024/3015/oj}{eur-lex.europa.eu
--- Reg (EU) 2024/3015} \\
\end{longtable}
}

\textbf{Objective.} Prohibits placing or making available on the EU
market --- or exporting --- any product made with forced labour at any
stage of its production, with investigation and enforcement powers for
authorities.

\textbf{In the ecosystem.} Adds enforcement teeth to forced-labour risk
in fishing and processing supply chains; documented due diligence
reduces investigation exposure and reinforces S2 value-chain work.

\textbf{Status flag.} \emph{Applies from 14 Dec 2027 --- it is a
future obligation, not yet enforceable. State the date when citing it.}

General Food Law --- General principles and requirements of food law

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & EU (establishes EFSA and the RASFF alert system) \\
\textbf{Reference / founded} & Regulation (EC) No 178/2002 \\
\textbf{Status · role} & EU regulation; directly applicable · Asker \\
\textbf{Stakeholder source} & Regulator (food safety) \\
\textbf{Official source} &
\href{https://eur-lex.europa.eu/eli/reg/2002/178/oj}{eur-lex.europa.eu
--- Reg (EC) 178/2002} \\
\end{longtable}
}

\textbf{Objective.} Lays down general food-safety principles, the
\textquotesingle one-step-back / one-step-forward\textquotesingle{}
traceability duty, and the Rapid Alert System for Food and Feed (RASFF).

\textbf{In the ecosystem.} Tier-1 food-integrity baseline; underpins
food-safety incident, recall and traceability topics that sit beneath
both regulatory and customer (GFSI) demands.

OSH Framework Directive --- Safety and health of workers at work

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & EU \\
\textbf{Reference / founded} & Directive 89/391/EEC \\
\textbf{Status · role} & EU directive; transposed nationally · Asker \\
\textbf{Stakeholder source} & Regulator (horizontal, social) \\
\textbf{Official source} &
\href{https://eur-lex.europa.eu/eli/dir/1989/391/oj}{eur-lex.europa.eu
--- Dir 89/391/EEC} \\
\end{longtable}
}

\textbf{Objective.} Establishes the employer\textquotesingle s general
duty to ensure workers\textquotesingle{} health and safety --- risk
assessment, prevention principles, information and training.

\textbf{In the ecosystem.} Legal baseline behind own-workforce
health-and-safety (S1) disclosure; especially salient for cutting lines,
machinery and cold stores in seafood processing.

Working Time Directive --- Organisation of working time

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & EU \\
\textbf{Reference / founded} & Directive 2003/88/EC \\
\textbf{Status · role} & EU directive; transposed nationally · Asker \\
\textbf{Stakeholder source} & Regulator (horizontal, social) \\
\textbf{Official source} &
\href{https://eur-lex.europa.eu/eli/dir/2003/88/oj}{eur-lex.europa.eu
--- Dir 2003/88/EC} \\
\end{longtable}
}

\textbf{Objective.} Sets minimum requirements for working hours, rest
periods, breaks, paid annual leave and night work.

\textbf{In the ecosystem.} Underpins the working-conditions elements of
own-workforce (S1) disclosure, but its mandated datapoints --- working
hours, rest periods and paid annual leave --- fall outside the
cohort-defined topic universe; its demand is therefore recorded in the
out-of-frame register rather than scored on an in-universe row.

F-gas Regulation --- Fluorinated greenhouse gases

{\def\LTcaptype{none}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & EU (Parliament \& Council) \\
\textbf{Reference / founded} & Regulation (EU) 2024/573 --- in force March 2024 (repealing Reg 517/2014) \\
\textbf{Status · role} & EU regulation; mandatory, directly applicable · Asker \\
\textbf{Stakeholder source} & Regulator (horizontal) \\
\textbf{Official source} & \href{https://eur-lex.europa.eu/eli/reg/2024/573/oj}{eur-lex.europa.eu --- Reg (EU) 2024/573} \\
\end{longtable}
}

\textbf{Objective.} Controls emissions of fluorinated greenhouse gases through an HFC phase-down (quota system), bans on high-GWP equipment, and direct operator duties: leak checks on equipment above 5 tonnes CO\textsubscript{2}-equivalent, leak-detection systems, recovery and certified handling, five-year record-keeping, and reporting via the F-gas portal (Art.~26).

\textbf{In the ecosystem.} Scope-conditional but acutely sector-relevant: refrigeration and freezing are core to seafood processing and the cold chain. It binds the processor where \emph{fluorinated} (HFC/HFO) refrigerants are used; firms on natural refrigerants (ammonia, CO\textsubscript{2}) fall outside its scope. Its refrigerant-emission data feed Scope~1 GHG (E1).

\textbf{Status flag.} \emph{Applies only to fluorinated-gas users --- governed in Phase~3 by a firm-profile include-switch.}

Whistleblowing Directive --- Protection of persons reporting breaches of EU law

{\def\LTcaptype{none}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & EU (Parliament \& Council) \\
\textbf{Reference / founded} & Directive (EU) 2019/1937 --- in force for 50--249-employee entities since 17 Dec 2023 \\
\textbf{Status · role} & EU directive; transposed nationally; mandatory · Asker \\
\textbf{Stakeholder source} & Regulator (horizontal) \\
\textbf{Official source} & \href{https://eur-lex.europa.eu/eli/dir/2019/1937/oj}{eur-lex.europa.eu --- Dir (EU) 2019/1937} \\
\end{longtable}
}

\textbf{Objective.} Requires private legal entities with 50 or more employees to establish confidential \emph{internal reporting channels} and procedures, protect reporters from retaliation, give feedback within set timeframes, and keep records of reports.

\textbf{In the ecosystem.} Binds the processor directly as an employer. It is admitted on the same basis as the OSH Framework Directive --- a binding organisation-level governance \emph{mechanism} on a G/S topic, not a public-disclosure mandate. Its content maps to business-conduct (G1) and the worker grievance / raise-a-concern channels in S1-3 and S2-3.

Pay Transparency Directive --- Equal pay and pay transparency

{\def\LTcaptype{none}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & EU (Parliament \& Council) \\
\textbf{Reference / founded} & Directive (EU) 2023/970 --- transposition deadline 7 June 2026 \\
\textbf{Status · role} & EU directive; phased by headcount; mandatory · Asker \\
\textbf{Stakeholder source} & Regulator (horizontal) \\
\textbf{Official source} & \href{https://eur-lex.europa.eu/eli/dir/2023/970/oj}{eur-lex.europa.eu --- Dir (EU) 2023/970} \\
\end{longtable}
}

\textbf{Objective.} Strengthens equal-pay enforcement and introduces mandatory \emph{gender pay-gap reporting}: employers with 250+ employees report annually; 150--249 every three years (first report 2027 on 2026 data); 100--149 from 2031. Triggers a joint pay assessment where an unjustified gap of 5\% or more is found.

\textbf{In the ecosystem.} Binds the processor directly above the headcount trigger. Its mandated datapoint (the gender pay gap) corresponds to ESRS~S1-16, which is \emph{outside} the cohort-defined topic universe --- so its demand is recorded in the out-of-frame register rather than scored on an in-universe row.

\textbf{Status flag.} \emph{Applies at or above 150 employees --- governed in Phase~3 by a firm-profile include-switch.}

VMP Regulation --- Veterinary Medicinal Products Regulation

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & EU \\
\textbf{Reference / founded} & Regulation (EU) 2019/6 --- applies from
2022 \\
\textbf{Status · role} & EU regulation; directly applicable · Asker \\
\textbf{Stakeholder source} & Regulator (sector) \\
\textbf{Official source} &
\href{https://eur-lex.europa.eu/eli/reg/2019/6/oj}{eur-lex.europa.eu ---
Reg (EU) 2019/6} \\
\end{longtable}
}

\textbf{Objective.} Governs authorisation and use of veterinary
medicines, including measures to curb antimicrobial use and resistance
(e.g. restrictions on prophylactic and metaphylactic antibiotics).

\textbf{In the ecosystem.} The regulatory anchor for the aquaculture
antimicrobial-use (AMU) topic, alongside the Farm to Fork
antimicrobial-reduction target --- one of the five structural-gap topics
in the matrix.

1C · Customer-driven standards \& audits

MSC --- Marine Stewardship Council

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & Independent non-profit (founded 1997 by Unilever \&
WWF; fully independent from 1999), London \\
\textbf{Reference / founded} & Founded 1997 \\
\textbf{Status · role} & Voluntary certification --- Fisheries Standard
+ Chain of Custody · Asker \\
\textbf{Stakeholder source} & Customer \\
\textbf{Official source} & \href{https://www.msc.org}{msc.org} \\
\end{longtable}
}

\textbf{Objective.} Certifies wild-capture fisheries against a
science-based sustainability standard (stock status, ecosystem impact,
management) and tracks certified product through the supply chain via a
Chain-of-Custody standard, assessed by accredited third-party
conformity-assessment bodies.

\textbf{In the ecosystem.} A de-facto market licence for many
wild-capture retail listings; its Chain of Custody sits squarely on the
traceability corridor. Wild-caught scope only.
\emph{Performance/outcome-based}: certifies measured stock and ecosystem
results against science-based benchmarks, not merely good practice. The
processor holds the \emph{Chain of Custody} certification (traceability
and segregation); the Fisheries Standard (Principles 1--3) certifies the
\emph{fishery}, not the processor --- so the processor's direct MSC
obligation is traceability, while stock and ecosystem performance reach
it as a sourcing requirement (direct on CoC, indirect on P1/P2).

\textbf{Evidence.} Peer-reviewed work links certification to healthier
stock indicators --- Gutiérrez et al. (2012, PLoS ONE) found the
ecolabel conveys reliable stock-health information --- though
effect-size and selection-bias debates persist (see
\textquotesingle Shifting focus: the impacts of sustainable seafood
certification\textquotesingle, PLoS ONE). Use as associative, not
causal, evidence.

ASC --- Aquaculture Stewardship Council

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & Non-profit (founded 2010 by WWF \& IDH), Utrecht \\
\textbf{Reference / founded} & Founded 2010 \\
\textbf{Status · role} & Voluntary certification --- species standards +
Chain of Custody · Asker \\
\textbf{Stakeholder source} & Customer \\
\textbf{Official source} & \href{https://asc-aqua.org}{asc-aqua.org} \\
\end{longtable}
}

\textbf{Objective.} Certifies responsibly farmed seafood against
species-specific environmental and social standards (water quality,
feed, disease/antimicrobials, labour), with a chain-of-custody linked to
the MSC CoC programme.

\textbf{In the ecosystem.} Market expectation for farmed seafood such as
salmon, shrimp and pangasius; covers welfare, antimicrobials, water and
feed. Aquaculture scope only. \emph{Performance/outcome-based}, with
measurable environmental and social metric thresholds (effluent, FFDR,
mortality, labour) --- distinct from GLOBALG.A.P.'s good-practice
compliance model.

\textbf{Evidence.} ASC and MSC merged their chain-of-custody programme;
for farmed-stock outcome evidence the literature is thinner and more
recent than for wild MSC --- flag any effectiveness claim as emerging.
{[}LIKELY{]}

BAP --- Best Aquaculture Practices

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & Global Seafood Alliance (formerly Global
Aquaculture Alliance), USA \\
\textbf{Reference / founded} & From 2002 onwards \\
\textbf{Status · role} & Voluntary certification --- multi-star,
full-chain · Asker \\
\textbf{Stakeholder source} & Customer \\
\textbf{Official source} &
\href{https://www.bapcertification.org}{bapcertification.org} \\
\end{longtable}
}

\textbf{Objective.} Certifies aquaculture facilities across the chain
--- hatchery, farm, feed mill and processing plant --- against
environmental, social, food-safety and animal-welfare criteria, using a
\textquotesingle star\textquotesingle{} rating for chain completeness.

\textbf{In the ecosystem.} Alternative or complement to ASC and the only
major scheme certifying feed mills directly; strong in North American
buyer requirements. Aquaculture scope.

\textbf{Versus GLOBALG.A.P. and ASC.} Like GLOBALG.A.P. it is a
good-practice / compliance scheme (food-safety-rooted, GFSI-recognised)
rather than an outcome-based one like MSC/ASC. But unlike farm-only
GLOBALG.A.P., BAP's star system certifies the \emph{processing plant}
itself, so a processor can hold BAP on its own facility --- a direct
certification, not merely a sourcing or chain-of-custody link.

\textbf{Status flag.} \emph{The Global Seafood Alliance also runs Best
Seafood Practices (BSP), the wild-capture counterpart to BAP. BSP is
young (launched 2022) and immature in uptake; it is not scored, noted
here for completeness.}

GLOBALG.A.P. --- Good Aquaculture Practice (Aquaculture standard)

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & FoodPLUS GmbH (origin EUREPGAP, 1997), Germany \\
\textbf{Reference / founded} & 1997 (EUREPGAP) / aquaculture module
2007 \\
\textbf{Status · role} & Voluntary B2B certification · Asker \\
\textbf{Stakeholder source} & Customer \\
\textbf{Official source} &
\href{https://www.globalgap.org}{globalgap.org} \\
\end{longtable}
}

\textbf{Objective.} Business-to-business good-practice certification
covering food safety, environmental management and worker welfare at the
production stage, with an aquaculture-specific module.

\textbf{In the ecosystem.} A common retailer prerequisite, particularly
for farmed produce; often the production-stage layer beneath a
retailer\textquotesingle s social and food-safety asks. Aquaculture
scope (for the module in view).

\textbf{Versus MSC/ASC.} GLOBALG.A.P. certifies \emph{conformity with
good-practice control points} --- farm management and food safety (it is
the only aquaculture farming standard recognised by GFSI) --- rather than
environmental \emph{performance outcomes}. Where MSC and ASC assess
measured results against science-based stock, ecosystem and effluent
benchmarks and carry a consumer ecolabel, GLOBALG.A.P. is a B2B
compliance assurance with no consumer-facing claim. Its newest version is
shifting toward a more outcome-based principles-and-criteria structure ---
verify which version a given firm holds.

\textbf{Note.} Scored at processor level via the sourcing requirement and
its Chain of Custody (the same channel as MSC/ASC). It overlaps BAP ---
both are good-practice aquaculture certifications --- so the two form a
substitution / dedup cluster: apply an independence weight so the
good-practice signal is not double-counted.

Friend of the Sea --- Friend of the Sea

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & World Sustainability Organization (founded 2008),
Italy \\
\textbf{Reference / founded} & Founded 2008 \\
\textbf{Status · role} & Voluntary certification (wild + farmed) ·
Asker \\
\textbf{Stakeholder source} & Customer \\
\textbf{Official source} &
\href{https://friendofthesea.org}{friendofthesea.org} \\
\end{longtable}
}

\textbf{Objective.} Certifies wild and farmed seafood (and other
products) against sustainability criteria, with an ecolabel.

\textbf{In the ecosystem.} An additional ecolabel option with lower
retail penetration than MSC/ASC but present in the landscape; covers
both production systems.

\textbf{Note.} \emph{Kept in the scored matrix --- a processor can hold
its Chain-of-Custody / product certification. But it overlaps MSC/ASC (a
product + traceability ecolabel covering both production systems) and has
low retail uptake, so it is largely a redundant echo of the dominant
ecolabels: apply a low independence weight in the dependency-adjusted
aggregation so its signal is not double-counted.}

IFS Food --- International Featured Standards --- Food

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & IFS Management GmbH (German/French/Italian retail
origin) \\
\textbf{Reference / founded} & Founded 2003 \\
\textbf{Status · role} & Voluntary, GFSI-benchmarked food-safety
certification · Asker \\
\textbf{Stakeholder source} & Customer \\
\textbf{Official source} &
\href{https://www.ifs-certification.com}{ifs-certification.com} \\
\end{longtable}
}

\textbf{Objective.} Certifies food-safety and quality-management systems
for processors supplying (mainly continental-European) retail.

\textbf{In the ecosystem.} Tier-2 market licence --- effectively
non-negotiable for retail listing; its supplier-control requirements
anchor the supplier-engagement corridor.

BRCGS --- Brand Reputation Compliance Global Standards

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & Owned by LGC (formerly British Retail
Consortium) \\
\textbf{Reference / founded} & From 1998 \\
\textbf{Status · role} & Voluntary, GFSI-benchmarked food-safety
certification · Asker \\
\textbf{Stakeholder source} & Customer \\
\textbf{Official source} & \href{https://www.brcgs.com}{brcgs.com} \\
\end{longtable}
}

\textbf{Objective.} Global food-safety and quality certification widely
required by UK and international retailers.

\textbf{In the ecosystem.} A near-substitute for IFS Food with the same
market-licence function; buyer choice between them is largely
geographic.

FSSC 22000 --- Food Safety System Certification 22000

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & Foundation FSSC \\
\textbf{Reference / founded} & Founded 2009 (built on ISO 22000) \\
\textbf{Status · role} & Voluntary, GFSI-benchmarked · Asker \\
\textbf{Stakeholder source} & Customer \\
\textbf{Official source} & \href{https://www.fssc.com}{fssc.com} \\
\end{longtable}
}

\textbf{Objective.} Food-safety system certification combining ISO 22000
with sector-specific prerequisite programmes.

\textbf{In the ecosystem.} The third common GFSI-benchmarked option;
interchangeable with IFS/BRCGS in most buyer specifications.

Sedex / SMETA --- Supplier Ethical Data Exchange / Sedex Members Ethical
Trade Audit

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & Sedex --- membership platform; SMETA is its audit
methodology, UK \\
\textbf{Reference / founded} & Founded 2004 \\
\textbf{Status · role} & Voluntary social-audit platform / methodology ·
Asker \\
\textbf{Stakeholder source} & Customer \\
\textbf{Official source} & \href{https://www.sedex.com}{sedex.com} \\
\end{longtable}
}

\textbf{Objective.} Two linked instruments. Members complete a
\emph{Self-Assessment Questionnaire} (SAQ) --- a broad self-disclosure
across five categories (labour, health \& safety, environment, business
ethics, management) --- and undergo \emph{SMETA}, a risk-based
third-party social audit that verifies and follows up on the SAQ. The SAQ
sets the breadth of topics disclosed; SMETA is the verification layer.

\textbf{In the ecosystem.} A common retailer-required disclosure-plus-audit.
It is scored on the \emph{SAQ}'s breadth --- it self-discloses across
E/S/G, deepest on labour and health \& safety (where SMETA also audits, so
those score 2) and lighter on environment and business ethics (score 1)
--- with SMETA as the risk-based verification that follows up on flagged
risks. Direct: the processor completes its own SAQ and is audited at its
own site.

EcoVadis --- EcoVadis

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & EcoVadis SAS --- private ratings company (founded
2007), France \\
\textbf{Reference / founded} & Founded 2007 \\
\textbf{Status · role} & Voluntary sustainability rating / assessment ·
Asker \\
\textbf{Stakeholder source} & Customer \\
\textbf{Official source} & \href{https://ecovadis.com}{ecovadis.com} \\
\end{longtable}
}

\textbf{Objective.} Scores suppliers on environment, labour \& human
rights, ethics and sustainable procurement through a documentation-based
assessment yielding a medal/percentile scorecard.

\textbf{In the ecosystem.} Frequently requested by large retail and B2B
buyers; its broad multi-topic scope makes it a strong hotspot driver in
the demand matrix.

1D · Investor-driven frameworks

CDP --- CDP (formerly Carbon Disclosure Project)

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & CDP --- non-profit, UK/global \\
\textbf{Reference / founded} & Founded 2000 \\
\textbf{Status · role} & Voluntary disclosure system (investor- and
buyer-requested) · Asker \\
\textbf{Stakeholder source} & Customer --- the dominant transmitter for a
B2B seafood SME is the retailer's CDP Supply Chain request rather than the
capital market \\
\textbf{Official source} & \href{https://www.cdp.net}{cdp.net} \\
\end{longtable}
}

\textbf{Objective.} Runs a global environmental disclosure system
(climate, water security, forests) through scored questionnaires (A--D)
used by investors and large buyers to request standardised data.

\textbf{Two routes to the SME.} CDP reaches a seafood SME through two
distinct mechanisms. (i) \emph{CDP Supply Chain}: a large retailer or
buyer that is itself a CDP member requests its suppliers --- including the
SME --- to disclose through the full CDP questionnaire, so the demand
arrives as a customer cascade. (ii) \emph{CDP's dedicated SME
questionnaire}: a lighter, proportionate version that the SME completes
directly. The two routes differ in depth and in who initiates them, but
both fix the same standardised, datapoint-defined environmental fields,
which is why the column can reach 2 on the datapoints CDP specifically
requires.

\textbf{In the ecosystem.} A common investor and large-customer climate
\& water questionnaire; strong overlap with GHG and water topics, and a
frequent conduit by which buyer demand reaches the SME.

SBTi --- Science Based Targets initiative

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & Founded by CDP, UN Global Compact, WRI and WWF (now
an incorporated entity) \\
\textbf{Reference / founded} & Founded 2015 \\
\textbf{Status · role} & Voluntary target-validation · Asker \\
\textbf{Stakeholder source} & Investor \\
\textbf{Official source} &
\href{https://sciencebasedtargets.org}{sciencebasedtargets.org} \\
\end{longtable}
}

\textbf{Objective.} Defines and independently validates corporate
emissions-reduction targets aligned with climate science (1.5°C
pathways), including FLAG (Forest, Land and Agriculture) guidance
relevant to feed and protein.

\textbf{In the ecosystem.} Signals a proactive climate posture;
validated targets feed CDP and ESRS E1 climate disclosure.

TCFD --- Task Force on Climate-related Financial Disclosures

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & Established by the Financial Stability Board;
monitoring transferred to the ISSB in 2023 \\
\textbf{Reference / founded} & Created 2015; recommendations 2017 \\
\textbf{Status · role} & Voluntary recommendations --- now largely
absorbed into IFRS S2 · Asker / Enabler \\
\textbf{Stakeholder source} & Investor \\
\textbf{Official source} &
\href{https://www.fsb-tcfd.org}{fsb-tcfd.org} \\
\end{longtable}
}

\textbf{Objective.} Provided a four-pillar framework --- governance,
strategy, risk management, and metrics \& targets --- for
climate-related financial disclosure.

\textbf{In the ecosystem.} The conceptual backbone now carried forward
by IFRS S2 and ESRS E1; the standalone task force was disbanded after
2023, so treat it as legacy architecture rather than a live ask.

IFRS S1 / S2 --- ISSB Sustainability \& Climate Disclosure Standards

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & International Sustainability Standards Board, under
the IFRS Foundation (created at COP26, 2021) \\
\textbf{Reference / founded} & Standards issued 2023 \\
\textbf{Status · role} & Voluntary baseline; becoming mandatory via
national adoption · Asker \\
\textbf{Stakeholder source} & Investor \\
\textbf{Official source} &
\href{https://www.ifrs.org/groups/international-sustainability-standards-board/}{ifrs.org
--- ISSB} \\
\end{longtable}
}

\textbf{Objective.} A global baseline of investor-focused disclosure: S1
(general sustainability-related financial information) and S2 (climate),
on a financial-materiality (\textquotesingle outside-in\textquotesingle)
basis.

\textbf{In the ecosystem.} The capital-market counterpart to ESRS; most
relevant where a seafood firm has an internationally listed parent or
investors in adopting jurisdictions.

\textbf{Coding note.} IFRS~S1 incorporates the SASB Standards as the
source of industry-specific metrics, but SASB has no dedicated seafood
standard --- a seafood processor falls under the generic
\emph{Processed Foods} industry, with no seafood-specific datapoints.
The column is therefore scored on its activity-agnostic climate core
(S2: GHG, targets, transition plan), not on sector metrics, and reaches
the sub-threshold SME indirectly through investor data requests. This
also distinguishes it from the EU Taxonomy: IFRS imposes defined
datapoints regardless of activity, so it is scored, whereas the
Taxonomy's activity-eligibility gate returns null for seafood processing
and is not.

SASB Standards --- Sustainability Accounting Standards Board Standards

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & Now under the ISSB / IFRS Foundation (via the 2022
Value Reporting Foundation merger) \\
\textbf{Reference / founded} & Codified 2018 \\
\textbf{Status · role} & Voluntary industry-specific metrics · Asker /
Enabler \\
\textbf{Stakeholder source} & Investor \\
\textbf{Official source} &
\href{https://sasb.ifrs.org}{sasb.ifrs.org} \\
\end{longtable}
}

\textbf{Objective.} Industry-specific, financially-material
sustainability metrics, including an agriculture/protein lens.

\textbf{In the ecosystem.} Feeds investor metric expectations and now
underpins the industry guidance within the ISSB standards rather than
operating as a separate ask.

Coller FAIRR --- Coller FAIRR Protein Producer Index

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & FAIRR Initiative --- investor network founded by
Jeremy Coller (2015) \\
\textbf{Reference / founded} & Index since 2018 \\
\textbf{Status · role} & Voluntary investor benchmark · Asker \\
\textbf{Stakeholder source} & Investor \\
\textbf{Official source} & \href{https://www.fairr.org}{fairr.org} \\
\end{longtable}
}

\textbf{Objective.} Ranks the largest listed animal-protein producers
(including aquaculture) on ESG risk factors --- feed, antibiotics,
welfare, emissions, deforestation --- for investor use.

\textbf{In the ecosystem.} Sector-specific investor scrutiny that a firm
is rated by rather than a member of; directly relevant for large/listed
producers, largely out of scope for an unlisted 50--250-employee SME.

DJSI --- Dow Jones Sustainability Indices

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & S\&P Global (Corporate Sustainability
Assessment) \\
\textbf{Reference / founded} & Since 1999 \\
\textbf{Status · role} & Voluntary investor index / assessment ·
Asker \\
\textbf{Stakeholder source} & Investor \\
\textbf{Official source} &
\href{https://www.spglobal.com/esg/csa}{spglobal.com --- CSA} \\
\end{longtable}
}

\textbf{Objective.} Selects index constituents via an annual Corporate
Sustainability Assessment scoring ESG performance.

\textbf{In the ecosystem.} Relevant mainly to large listed parents;
signals investor-grade ESG performance and is, like FAIRR, a rating-of
rather than a membership.

TNFD --- Taskforce on Nature-related Financial Disclosures

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & Market-led, government-backed taskforce (launched
2021) \\
\textbf{Reference / founded} & Recommendations finalised 2023 \\
\textbf{Status · role} & Voluntary recommendations · Asker / Enabler \\
\textbf{Stakeholder source} & Investor \\
\textbf{Official source} & \href{https://tnfd.global}{tnfd.global} \\
\end{longtable}
}

\textbf{Objective.} A framework (the LEAP approach) for assessing and
disclosing nature- and biodiversity-related dependencies, impacts, risks
and opportunities.

\textbf{In the ecosystem.} The emerging biodiversity/ecosystem lens,
highly relevant to marine impacts and dependencies (E4); an EFRAG--TNFD
correspondence mapping links it to ESRS.

1E · Industry / NGO voluntary frameworks

GRI --- GRI Standards (incl. GRI 13)

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & Global Reporting Initiative --- independent
standards body (founded 1997), Amsterdam \\
\textbf{Reference / founded} & GRI 13 \textquotesingle Agriculture,
Aquaculture and Fishing Sectors\textquotesingle, 2022 \\
\textbf{Status · role} & Voluntary reporting standards · Asker \\
\textbf{Stakeholder source} & Industry / NGO \\
\textbf{Official source} &
\href{https://www.globalreporting.org}{globalreporting.org} \\
\end{longtable}
}

\textbf{Objective.} The most widely used global sustainability-reporting
standards, on an impact (\textquotesingle inside-out\textquotesingle)
materiality basis; GRI 13 enumerates the topics likely material to the
agriculture, aquaculture and fishing sector.

\textbf{In the ecosystem.} The sector-materiality benchmark used in this
thesis\textquotesingle s topic-universe cross-check and the single most
seafood-relevant existing disclosure map; an EFRAG--GRI interoperability
index (Nov 2024) aligns GRI with ESRS.

UN Global Compact --- United Nations Global Compact

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & United Nations \\
\textbf{Reference / founded} & Founded 2000 \\
\textbf{Status · role} & Voluntary corporate-citizenship initiative ·
Asker \\
\textbf{Stakeholder source} & Industry / NGO \\
\textbf{Official source} &
\href{https://unglobalcompact.org}{unglobalcompact.org} \\
\end{longtable}
}

\textbf{Objective.} Commits signatories to ten principles on human
rights, labour, environment and anti-corruption, with an annual
Communication on Progress.

\textbf{In the ecosystem.} A common voluntary commitment and normative
reference for governance and social topics; entry-level credibility
signal rather than a data standard.

Move to Minus 15 --- Move to -15 °C Coalition

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & Industry coalition (convened with DP World and
supply-chain partners) \\
\textbf{Reference / founded} & Initiated 2023 \\
\textbf{Status · role} & Voluntary initiative · Asker \\
\textbf{Stakeholder source} & Industry / NGO \\
\textbf{Official source} &
\href{https://movetominus15.com}{movetominus15.com} \\
\end{longtable}
}

\textbf{Objective.} A coalition exploring raising the frozen-food
storage standard from -18 °C to -15 °C to cut energy use and emissions
without compromising safety.

\textbf{In the ecosystem.} A sector-decarbonisation initiative touching
energy (E1) and cold-chain operations; directly relevant to
frozen-seafood logistics.

\textbf{Status flag.} \emph{Excluded from the scored coverage matrix: as
a coalition committing members to \emph{explore} a -15 °C standard, it
sets no reporting, due-diligence or certification requirement and imposes
no own organisation-level obligation on the processor (fails Gate~2). It
is retained in the Phase~2 landscape as a Tier~4 cold-chain engagement
context. Governance and evidence base are still developing; treat
membership/standard details as {[}ASSUMPTION{]} and verify before citing.}

ISSF --- International Seafood Sustainability Foundation

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & Science-and-industry partnership (scientists, tuna
processors, WWF) \\
\textbf{Reference / founded} & Founded 2009 \\
\textbf{Status · role} & Voluntary (tuna-focused) · Asker \\
\textbf{Stakeholder source} & Industry / NGO \\
\textbf{Official source} &
\href{https://www.iss-foundation.org}{iss-foundation.org} \\
\end{longtable}
}

\textbf{Objective.} Promotes science-based tuna conservation through a
set of binding \emph{Conservation Measures} (33 as of 2024, covering
bycatch mitigation, transshipment oversight, vessel monitoring and
traceability) and the ProActive Vessel Register. Participating companies
commit to these measures and are independently audited each year by MRAG
Americas, with per-company compliance reports published (99.6\% conformance
in the 2024 audit cycle).

\textbf{In the ecosystem.} Tuna-specific; relevant only to processors that
handle tuna. For a participating tuna company the audited conservation
measures are a binding conformity mechanism (passing Gate~2, like a SMETA
audit), so ISSF is a genuine compliance framework rather than mere context.

\textbf{Coding note.} It is carried as a firm-profile \emph{opt-in}
column, switched off by default and activated only where the modelled SME
processes tuna. Its absence from the scored set for a generic seafood
processor is therefore a firm-profile trigger, not a Gate-2 exclusion ---
distinct from instruments dropped for imposing no own obligation (e.g.\ a
FIP). Wild-caught scope.

FIPs / FisheryProgress --- Fishery Improvement Projects /
FisheryProgress.org

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & FisheryProgress.org operated by FishChoice in
partnership with the Sustainable Fisheries Partnership (SFP) \\
\textbf{Reference / founded} & Platform from \textasciitilde2017 \\
\textbf{Status · role} & Voluntary improvement mechanism (not a
certification) · Asker \\
\textbf{Stakeholder source} & Industry / NGO \\
\textbf{Official source} &
\href{https://fisheryprogress.org}{fisheryprogress.org} \\
\end{longtable}
}

\textbf{Objective.} Time-bound, multi-stakeholder projects that move a
fishery toward sustainability, with public progress tracking on a
central platform; many are engineered toward an eventual MSC assessment.

\textbf{In the ecosystem.} A sourcing pathway for not-yet-certified
fisheries and buyer-recognised improvement evidence. A retailer may well
\emph{ask} a processor to source from a FIP, which is why the instrument
is retained at Tier~3 (competitive/market expectation) in the Phase~2
Necessity matrix. Crucially, however, that ask is an obligation to
\emph{engage and source}, not to report: the FisheryProgress reporting
requirements --- workplans, progress updates and the 2021 Human Rights
and Social Responsibility Policy (public commitment, vessel list,
grievance mechanism, forced-labour self-evaluation) --- bind the FIP
itself (typically NGO- or fishery-led), not the buying processor. A FIP
is pre-certification and carries no chain-of-custody certificate the
processor can hold, so unlike MSC or ASC it transmits no standard onto
the processor's own datapoints. Wild scope.

\textbf{Status flag.} \emph{Excluded from the scored coverage matrix
(Phase~3): the instrument fails Gate~2, imposing no own organisation-level
reporting, due-diligence or certification obligation on the processor ---
only engagement and sourcing. It is retained in the Phase~2 landscape and
scope log as a Tier~3 sourcing-engagement context, and the topics a FIP
would touch (wild-stock status, traceability) are already captured by the
MSC, ASC, GDST and IUU columns, so scoring it would only add redundant,
low-independence signal.}

\textbf{Evidence.} FIP effectiveness is contested in the literature:
reviews flag stalled progress and weak accountability once market access
is gained (\textquotesingle Shifting focus: the impacts of sustainable
seafood certification\textquotesingle, PLoS ONE). Present FIP
participation as intent, not outcome.

SeaBOS --- Seafood Business for Ocean Stewardship

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & Science--business initiative linked to the
Stockholm Resilience Centre \\
\textbf{Reference / founded} & Founded 2016 \\
\textbf{Status · role} & Voluntary CEO-level coalition · Asker \\
\textbf{Stakeholder source} & Industry / NGO \\
\textbf{Official source} & \href{https://seabos.org}{seabos.org} \\
\end{longtable}
}

\textbf{Objective.} Connects the largest
(\textquotesingle keystone\textquotesingle) seafood companies with
science to drive ocean stewardship on IUU fishing, traceability and
climate.

\textbf{In the ecosystem.} A large-player, agenda-setting initiative
rather than an SME-facing demand; useful as a direction-of-travel
reference, aspirational for a 50--250-employee firm.

RSPO --- Roundtable on Sustainable Palm Oil

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & RSPO --- multi-stakeholder non-profit,
Malaysia/Switzerland \\
\textbf{Reference / founded} & Founded 2004 \\
\textbf{Status · role} & Voluntary certification + supply-chain models ·
Asker \\
\textbf{Stakeholder source} & Industry / NGO \\
\textbf{Official source} & \href{https://rspo.org}{rspo.org} \\
\end{longtable}
}

\textbf{Objective.} Certifies sustainable palm oil along the supply
chain via production standards and chain-of-custody models.

\textbf{In the ecosystem.} Relevant where palm derivatives appear in
aquaculture feed or processing inputs; intersects with the EUDR
due-diligence obligation.

Layer 2 --- Enablement / methodology layer

These instruments underpin the demand-side layer. They supply accounting
methods, management-system processes, benchmarks or data formats. A firm
builds capability on them, but they do not themselves request a
disclosure --- which is why they are excluded from the demand count yet
central to the capability-corridor argument.

GHG Protocol --- Greenhouse Gas Protocol

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & World Resources Institute (WRI) \& WBCSD \\
\textbf{Reference / founded} & Corporate Standard, 2001 \\
\textbf{Status · role} & Accounting methodology (not certifiable) ·
Enabler \\
\textbf{Stakeholder source} & Cross-cutting \\
\textbf{Official source} &
\href{https://ghgprotocol.org}{ghgprotocol.org} \\
\end{longtable}
}

\textbf{Objective.} The dominant corporate GHG accounting and reporting
methodology, defining Scopes 1, 2 and 3 and the rules for
boundary-setting and calculation.

\textbf{In the ecosystem.} Underlies every climate disclosure (ESRS E1,
CDP, SBTi); it does not itself demand reporting but makes emissions data
comparable --- a pure capability enabler.

ISO 14001 --- Environmental Management Systems

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & International Organization for Standardization \\
\textbf{Reference / founded} & 1996 / 2015 revision \\
\textbf{Status · role} & Management-system standard (certifiable) ·
Enabler \\
\textbf{Stakeholder source} & Cross-cutting \\
\textbf{Official source} &
\href{https://www.iso.org/iso-14001-environmental-management.html}{iso.org
--- ISO 14001} \\
\end{longtable}
}

\textbf{Objective.} Specifies requirements for an environmental
management system (EMS) on a plan-do-check-act basis.

\textbf{In the ecosystem.} Process backbone for environmental data;
supports E-topic capability without itself being a disclosure asker.

ISO 45001 --- Occupational Health \& Safety Management Systems

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & International Organization for Standardization \\
\textbf{Reference / founded} & 2018 \\
\textbf{Status · role} & Management-system standard (certifiable) ·
Enabler \\
\textbf{Stakeholder source} & Cross-cutting \\
\textbf{Official source} &
\href{https://www.iso.org/iso-45001-occupational-health-and-safety.html}{iso.org
--- ISO 45001} \\
\end{longtable}
}

\textbf{Objective.} Specifies requirements for an occupational health \&
safety management system.

\textbf{In the ecosystem.} Underpins own-workforce H\&S (S1) data and
processes; the management-system counterpart to the OSH
Directive\textquotesingle s legal duty.

ISO 50001 --- Energy Management Systems

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & International Organization for Standardization \\
\textbf{Reference / founded} & 2011 / 2018 revision \\
\textbf{Status · role} & Management-system standard (certifiable) ·
Enabler \\
\textbf{Stakeholder source} & Cross-cutting \\
\textbf{Official source} &
\href{https://www.iso.org/iso-50001-energy-management.html}{iso.org ---
ISO 50001} \\
\end{longtable}
}

\textbf{Objective.} Specifies requirements for an energy management
system, including energy-performance monitoring.

\textbf{In the ecosystem.} Supports the quality and auditability of
energy-consumption data (ESRS E1).

GSSI --- Global Sustainable Seafood Initiative

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & GSSI --- public-private partnership \\
\textbf{Reference / founded} & Founded 2015 \\
\textbf{Status · role} & Benchmarking tool (meta-standard) · Enabler \\
\textbf{Stakeholder source} & Customer (meta) \\
\textbf{Official source} &
\href{https://www.ourgssi.org}{ourgssi.org} \\
\end{longtable}
}

\textbf{Objective.} Benchmarks seafood certification schemes against
FAO-based criteria so buyers can recognise equivalence between
recognised schemes.

\textbf{In the ecosystem.} Reduces certification fragmentation --- it
tells buyers which schemes (MSC, ASC, BAP, GLOBALG.A.P.) it recognises
rather than demanding data itself; this is why it is treated as a
meta-layer and collapsed, not counted as an independent demand.

GFSI --- Global Food Safety Initiative

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & The Consumer Goods Forum \\
\textbf{Reference / founded} & Founded 2000 \\
\textbf{Status · role} & Benchmarking tool (meta-standard) · Enabler \\
\textbf{Stakeholder source} & Customer (meta) \\
\textbf{Official source} & \href{https://mygfsi.com}{mygfsi.com} \\
\end{longtable}
}

\textbf{Objective.} Benchmarks food-safety schemes (IFS, BRCGS, FSSC
22000) for mutual recognition --- \textquotesingle certified once,
accepted everywhere\textquotesingle.

\textbf{In the ecosystem.} The reason IFS, BRCGS and FSSC behave as
near-substitutes: buyers accept any GFSI-benchmarked scheme, so they
should not be triple-counted as independent demands.

GDST --- Global Dialogue on Seafood Traceability

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & GDST --- non-profit foundation (industry-led, 60+
partners) \\
\textbf{Reference / founded} & Convened 2017; GDST 1.0 standard, 2020 \\
\textbf{Status · role} & Voluntary data / interoperability standard ·
Enabler \\
\textbf{Stakeholder source} & Industry / NGO --- coded by issuer rather
than transmitter because, unlike the MSC/ASC certifications, GDST adoption
is not yet a standing retailer requirement; it remains an industry-led
interoperability standard the processor may adopt proactively \\
\textbf{Official source} & \href{https://thegdst.org}{thegdst.org} \\
\end{longtable}
}

\textbf{Objective.} A four-pillar traceability standard built on GS1
EPCIS: it defines internationally-agreed \emph{Key Data Elements} (KDEs
--- the \emph{what}: species, catch/production area, gear, vessel,
dates, chain-of-custody events) attached to \emph{Critical Tracking
Events} (CTEs --- the \emph{when}), plus the technical specifications for
interoperable, business-to-business exchange of that data (the
\emph{how}).

\textbf{In the ecosystem.} The technical spine of the traceability
corridor: it lets traceability data flow across IUU/Control-Reg
requirements, MSC/ASC chain-of-custody and ESRS value-chain reporting.
Buyers may require it, giving it a dual asker/enabler character.

\textbf{Coding note.} Although often described as a data-format standard,
GDST is \emph{not} purely interoperability: its KDEs are content. A
GDST-conforming processor must itself capture and transmit those defined
traceability datapoints, so the instrument imposes an own
organisation-level data obligation (passing Gate~2) and is scored ---
narrowly --- on supply-chain traceability (2) and, via the catch-area
KDE, wild-stock sourcing transparency (1). This is the line that
separates it from a FIP, whose reporting binds the project, not the
processor.

UN SDGs --- UN Sustainable Development Goals

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.2382}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.7618}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Behind it} & United Nations (Agenda 2030) \\
\textbf{Reference / founded} & 2015 \\
\textbf{Status · role} & Normative goal framework (not a reporting
standard) · Enabler \\
\textbf{Stakeholder source} & Cross-cutting \\
\textbf{Official source} &
\href{https://sdgs.un.org/goals}{sdgs.un.org} \\
\end{longtable}
}

\textbf{Objective.} Seventeen goals and 169 targets providing a shared
global agenda for sustainable development; SDG 14 (\textquotesingle Life
below water\textquotesingle) is the most directly seafood-relevant.

\textbf{In the ecosystem.} Used by firms to frame and communicate
contribution; a reference frame rather than a data demand, which is why
it sits in the enabler layer.

\textbf{Status flag.} \emph{Excluded from the scored coverage matrix: the
SDGs are an aspirational goal framework with no corporate reporting
mechanism, no signatory commitment and no imposed datapoint --- a firm
maps its activities to them voluntarily. They impose no own
organisation-level obligation (fails Gate~2), distinguishing them from the
UN Global Compact, whose mandatory annual Communication on Progress is a
genuine reporting obligation and which is therefore retained and scored.
The SDGs are kept as cross-cutting framing context only.}

Selected literature \& evidence base

Sources cited in the evidence notes above, plus key peer-reviewed work
on seafood-certification effectiveness. URLs marked \textbf{(verified)}
were confirmed live during research; entries marked \textbf{{[}verify
citation{]}} should be checked for exact authorship, year and pagination
before inclusion in the bibliography.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Gutiérrez, N. L. et al. (2012). ``Eco-Label Conveys Reliable
  Information on Fish Stock Health to Seafood Consumers.'' PLoS ONE
  7(8): e43765.
  \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3424161/}{PMC3424161}
  \emph{--- (verified) Evidence that MSC-certified stocks trend
  healthier; associative, not causal.}
\item
  ``Shifting focus: the impacts of sustainable seafood certification.''
  PLoS ONE (open access).
  \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7239462/}{PMC7239462}
  \emph{--- (verified URL) Reviews certification outcomes and FIP
  accountability concerns. {[}verify citation{]} for authors/year.}
\item
  Sampson, G. S. et al. (2015). ``Secure sustainable seafood from
  developing countries.'' Science 348(6234): 504--506. \emph{---
  {[}verify citation{]} Widely cited on FIPs and market access for
  developing-country fisheries.}
\item
  EFRAG (2024). GRI--ESRS Interoperability Index; EFRAG--TNFD
  correspondence mapping (2024). \emph{--- Official interoperability
  mappings underpinning the framework-connection logic. Available via
  efrag.org.}
\item
  FAO. Guidelines for the Ecolabelling of Fish and Fishery Products from
  Marine Capture Fisheries (basis for GSSI benchmarking). \emph{---
  {[}verify citation{]} Primary normative reference behind GSSI
  recognition criteria.}
\end{enumerate}

\emph{Compiled June 2026. All official-source links and regulatory
statuses were current at the snapshot date; EU sustainability law (CSRD,
CSDDD, ESRS, EUDR) was under active simplification and should be
re-checked against EUR-Lex before citation.}
