% ===== RUBEN: streamlined rewrite of the Introduction. =====
% Storyline: many actors, seemingly different demands -> shared underlying
% data -> resource-limited firms don't build structured processes -> they
% stay reactive (lagging the wave) when they could be proactive (riding it)
% -> seafood processing as a revealing SME-dominated case -> three angles
% (cohort topics, ecosystem frameworks, topics x frameworks) -> the so-what.
% Your previous Introduction is preserved just above, inside \iffalse ... \fi.

\section{1. Introduction}\label{introduction}

%CL: ===== REVIEW PASS (relevant / coherent / so-what). My notes = %CL: ; add yours as %Li: and hand back. =====
%CL: OVERALL — the arc is strong and the so-what lands cleanly. Two cross-cutting things to settle before this goes to the companies:
%CL:  (1) NUMBER CONSISTENCY: the intro repeatedly says "four or five themes," but the analysis/tool build-first set is ~8 topics (energy E1-5, GHG E1-6, water E3-4, waste E5-5, biodiversity E4, health&safety S1-14, value-chain workers S2-1, traceability). Decide now how to reconcile: group the 8 into ~5 named theme-families here, so intro and Results tell the same story. A reviewer WILL notice if 1.2 promises five and Results shows eight.
%CL:  (2) MISSING CITATIONS: Adams & Abhayawansa, Christensen/Hail/Leuz, Ambrosetti, EFRAG, STECF, European Commission are all cited below but not yet in references.tex. Add to Zotero + re-export before send, or the bibliography breaks.
%CL: %Li:  <-- your overall reaction here

\subsection{1.1 A rising tide of sustainability
demands}\label{a-rising-tide-of-sustainability-demands} %LI: new title - because not really rising tide - has been there more like seeing the trees through the forsest ... 

Companies today face a steadily rising volume of requests for
sustainability information, and from a widening set of actors (%LI: source?)
Regulators legislate mandatory disclosures; retail and business customers
send sustainability questionnaires and require certifications as a
condition of supply; banks, investors and ratings agencies ask for climate
and ESG data before extending capital or a favourable score; and industry
coalitions and NGOs promote voluntary standards and commitments. Each of
these actors asks through its own channel, on its own timetable, and seeminly in its
own vocabulary. Seen one request at a time, the effect is
of a proliferating, uncoordinated wall of separate demands: the
\textquotesingle alphabet soup\textquotesingle{} of standards, frameworks
and ratings that has prompted sustained calls for harmonisation (Adams \&
Abhayawansa, 2022; Christensen, Hail \& Leuz, 2021). %LI: both sources are not correct - the first calls harmonisation a myth and the second source focusses on US - find appropriate sources. The bottleneck of the
sustainability transition, at the level of the wider economy, is
increasingly this fragmentation of instruments rather than any disagreement
about goals (The European House~--~Ambrosetti, 2026).

%LI - the problem is that though many of these initiatives are maybe indeed independent many come also from regulation - but different actors so you get cascading effects - should this be made clear already or leave as is?

For the firm on the receiving end, the practical experience is of a rising
tide: the volume and variety of what is asked keeps growing, and staying
above it seemingly demands more and more effort.

%CL: 1.1 — Relevant and coherent, good scene-setting. Only flag: the Ambrosetti (2026) "the bottleneck is fragmentation, not disagreement about goals" claim is doing real argumentative work. Make sure it's a citable, defensible source (title + page), because it's the sentence that justifies your whole premise and a company reader may push on it.
%CL: %Li:check comment above!

\subsection{1.2 Different questions, shared
answers}\label{different-questions-shared-answers}

The central observation motivating this thesis is that these demands are
far less independent than they appear. Although each actor phrases its
request differently, the \emph{underlying data} on which the requests draw
overlaps heavily %Li: source? A customer\textquotesingle s carbon-footprint question, an
investor\textquotesingle s climate metric and a regulatory emissions
disclosure all rest on the same greenhouse-gas accounting. A retailer's
traceability requirement, a certification\textquotesingle s chain-of-custody
audit, a due-diligence law and a value-chain disclosure standard all rest on
the same body of lot-level tracking data. Beneath a large number of
superficially different and fragmented questions sits a much smaller number of shared
datapoints, and these datapoints cluster into a handful of recurring
themes.

%LI: nuance - there are a lot of themes - arguably there are some central themes that give you the licence 'to operate' - should we manage this nuance?

This has a direct and practical implication. The dozens of instruments do
not resolve into dozens of separate data problems: they resolve into a
handful of recurring \emph{themes} --- climate and energy, water,
traceability, workforce and health and safety, business conduct --- and a
defined set of \emph{datapoints} beneath each. A firm that identifies those
few themes and learns how to produce their datapoints \emph{once}, in a
structured and reusable way, can answer most of what it is asked from a
single effort, rather than rebuilding an answer for every request. The task
is therefore not to respond to every actor separately; it is to find the
key themes on which the demands converge and to get good at the
data behind them %LI: to build processess once that can be reused for multiple actors - again, source?

%CL: 1.2 — This is the heart of your so-what and it reads well. COHERENCE CHECK: the theme list here (climate/energy, water, traceability, workforce + health&safety, business conduct) is FIVE and omits biodiversity (E4) and waste (E5-5) — both of which sit in your build-first set in Results. Either fold them in, or say explicitly that the headline five are the highest-convergence subset. This is the same number-consistency point from the overall note, and 1.2 is where you set the reader's expectation.
%CL: %Li:riding wave is maybe a bit strong - tone down please - more direct, les prozaic language! 
\subsection{1.3 Why resource-limited firms stay
reactive}\label{why-resource-limited-firms-stay-reactive}

The barrier to capturing that prize (%LI: explain again what that prize is in simple terms) is not intent but capacity. Small and
medium-sized enterprises (SMEs) operate with limited financial and
managerial resources, little dedicated reporting expertise and weak data
infrastructure (Battisti \& Perry, 2011). Under those constraints the
up-front investment in structured data processes --- the very investment
that would let a firm answer many frameworks from one dataset --- is rarely
made. Instead the firm responds \emph{reactively}: each new questionnaire,
audit or regulation is handled as a separate, one-off exercise, its data
assembled by hand and then set aside. Fragmented governance thus
pushes resource-constrained firms toward minimum-viable compliance rather
than strategic integration (Chan, Lai \& Kim, 2022).

The consequence is that every new demand arrives as a shock. Because no
reusable capability was built, each regulation or customer request triggers
a fresh scramble, and the firm is perpetually a step behind -  mostly
catching up to the last wave, rarely prepared for the next. This is the
distinction between a reactive and a proactive sustainability posture
(Sharma \& Vredenburg, 1998; Torugsa, O\textquotesingle Donohue \& Hecker,
2013), developed further in Chapter~3: the reactive firm pays the full,
isolated cost of each demand in turn, while the proactive firm, having
invested in shared capability, meets new demands at low marginal cost and
can turn compliance into competitive advantage. Two licences are at stake
at once: the \emph{legal} licence to operate, which direct regulation makes
non-negotiable, and the \emph{commercial} licence to sell, which large
customers and their financiers increasingly attach to the same
sustainability data. A firm that builds that data once secures both; a firm
that answers each demand in isolation risks either.

To extend the metaphor of the rising tide: the reactive firm expends its
energy simply not being swamped by each successive wave of demand. Yet the
same firm, given a clear view of \emph{where} the demands overlap and which
capabilities serve the most of them, could stop struggling against the
water and begin to ride it - using each compliance investment as a
platform for the next. Supplying that view is the purpose of this thesis:
to show, for a resource-limited firm, the few themes on which the many
demands converge --- and the datapoints behind them --- so that a handful of
well-chosen investments does the work of many.

%CL: 1.3 — Strong, sendable. The two-licences framing (legal licence to operate / commercial licence to sell) is exactly the business case the companies will care about. No change needed on substance; just confirm Battisti&Perry, Sharma&Vredenburg, Torugsa and Chan/Lai/Kim all resolve in references, and that the reactive/proactive promise "developed further in Chapter 3" is actually delivered in 3.1 (it is).
%CL: %Li:
\subsection{1.4 Seafood processing in Europe: a revealing
case}\label{seafood-processing-a-revealing-case}

To make the problem tractable and empirically grounded, this thesis studies
one sector in depth: European seafood processing. Three features make it a revealing case.

First, it is overwhelmingly a sector of SMEs. In 2023, only 66 of some
3{,}245 EU enterprises processing fish as their main activity, about two
per cent, employed 250 people or more, while micro-enterprises of ten
employees or fewer alone accounted for around two-thirds of all firms
(STECF, 2025). %LI: quote differently - first say what the amount of SMEs is and then hence only 2 percent non-SMEs - stronger
The great majority of the sector therefore sits squarely in the
resource-limited position described above.

Second, it carries a dense and fragmented governance load.
Seafood processors sit beneath the full stack of horizontal EU sustainability
law (the CSRD, CSDDD, the Packaging and Packaging Waste Regulation, the
Empowering Consumers Directive and more), a layer of
sector-specific regulation (the Common Fisheries Policy, the IUU and Control
Regulations, the EU Deforestation Regulation), \emph{and} a crowded field of
voluntary certifications, ratings and initiatives (MSC and ASC certification,
the GDST traceability standard, EcoVadis, CDP and others). Third, this
compounding pressure is not merely inferred from the regulatory map:
scoping interviews conducted for this research with small seafood processors
in Belgium and Germany confirmed it directly. Those firms described
simultaneous, overlapping requests from regulators, retail customers and
finance providers, and a response that was, for want of the capacity to
do otherwise, reactive and request-by-request.

Seafood is a case, however, not the point. The structure that makes it
revealing, many actors, overlapping underlying data, resource-limited
firms, and a reactive response for want of shared capability - recurs
across manufacturing and agri-food SME sectors facing the same widening set
of sustainability demands. Seafood processing is a sharp instance that
makes the general problem visible and testable; the diagnostic approach
developed here, and the method behind it, are intended to transfer.

%CL: 1.4 — Relevant and it's your empirical anchor, BUT the softest spot for a company audience: "scoping interviews conducted for this research with small seafood processors in Belgium and Germany confirmed it directly." The cohort firms reading this may ask who/how many/what method. Two options: (a) keep it as light context and soften to "informal scoping conversations," or (b) back it properly with N + method in the methodology so it's defensible. Also LOCK the STECF 25-15 figures (66 of ~3,245; ~2%; ~two-thirds micro) against the source before send. Decide (a) vs (b).
%CL: %Li: make informal scoping conversations! 

\subsection{1.5 The reporting frame and the SME
gap}\label{the-reporting-frame-and-the-sme-gap}

The European Union has made corporate sustainability reporting a central
lever of the Green Deal. The Corporate Sustainability Reporting Directive
(CSRD), implemented through the European Sustainability Reporting Standards
(ESRS), defines mandatory disclosures across environmental, social and
governance topics and introduces the double-materiality principle (EFRAG,
2021). But CSRD scope is restricted to large undertakings, and the Omnibus
simplification proposals narrow it further (European Commission, 2025); the
majority of European seafood processors fall below the thresholds and are
not directly in scope. To serve out-of-scope firms, EFRAG has developed the
Voluntary Sustainability Reporting Standard for SMEs (VSME), a proportionate
module reducing roughly a thousand ESRS datapoints to about a hundred.

Two gaps nonetheless remain. The step from no structured reporting at all to
even the basic VSME module is substantial for a small processor; and the
step from VSME to the fuller set of demands a firm actually faces in
practice --- customer questionnaires, certification audits, investor
requests --- is larger still. The reporting frame, in other words, tells an
SME what a proportionate disclosure \emph{looks like}, but not where, amid
its own particular thicket of overlapping demands, it should begin.

%CL: 1.5 — Coherent. Two accuracy checks: (i) "VSME ... roughly a thousand ESRS datapoints to about a hundred" — verify both counts, the ~1000 figure especially; (ii) attribution — VSME was DEVELOPED by EFRAG but RECOMMENDED by the European Commission (30 Jul 2025). Phrase so both are correct; right now it reads as EFRAG-only.
%CL: %Li: weird placing - either this has to be incorporated into previous paragraph or moved upfront - now it sits weirdly? maybe can be used in beginning to say where it started with the EU - cascading effects of this + requests of other information, etc? 
\subsection{1.6 Aim, approach and
contribution}\label{aim-approach-and-contribution}

This thesis addresses that question %LI: which question. Its aim is to identify, for a
resource-limited European seafood-processing SME, the key
disclosure themes - and the underlying datapoints - that, if built
first, satisfy the largest share of the overlapping demands the firm faces
across regulators, customers, investors and industry initiatives, and
thereby to support a shift from reactive compliance to proactive strategy. %LI: and ensure license to operate!

It approaches the problem from three complementary angles, rather than as a
single linear pipeline. The first angle establishes the \emph{ecosystem}:
the full landscape of instruments acting on the sector --- who demands what
--- organised by how obligatory each demand is and which stakeholder drives
it. %CL: 1.6 — Three-angles framing is clean and matches the restructure; the contribution paragraph is well-pitched. THIS is the natural home for your materiality-vs-IRO justification (task #17): one sentence, right after "common coordinate system," on why you analyse at the level of ESRS material TOPICS, not per-topic IROs (impacts/risks/opportunities) — because IROs are entity-specific and don't compare across the 31 instruments, whereas topics give the common grain a convergence mapping needs.
%CL: %Li: not IRO here - that's part of methodology. 

To compare and aggregate across this fragmented field, the analysis
adopts the mandatory European Sustainability Reporting Standards (ESRS) as a
common coordinate system %LI: Benchmark?, the most comprehensive and legally anchored
reference available. The second angle looks at \emph{topics through the
cohort}: benchmarked against that ESRS structure, what the
sector\textquotesingle s most advanced reporters actually disclose, which
establishes the realistic frontier of seafood disclosure and a universal
core of topics every mature firm reports. The third \emph{crosses the two}:
mapping which instruments demand which topics reveals where demand converges
--- the hotspots at which a single theme, once its datapoints are in place,
answers many actors at once --- and this cross-analysis is operationalised as a diagnostic tool
with which an individual SME can locate its own posture and identify the
topics it should build first.

The contribution is threefold. For seafood SMEs, the thesis offers a
diagnostic framework and a prioritised, sequenced set of topics tailored to a
firm\textquotesingle s situation. %LI: Add that is should be basis for strategic discussion within SMEs.
For the academic literature, it provides a
first systematic mapping of cross-stakeholder disclosure demand at sector
level and an empirically grounded application of the reactive/proactive
distinction. For EU policy, it supplies evidence to the ongoing
simplification debate (the Omnibus package, the VSME, the postponement of
sector standards) by identifying where the disclosure ecosystem converges,
and therefore where coordinating or capping demand would deliver
disproportionate reductions in SME burden. %LI: is this still acurate actually? The practical payoff is direct:
instead of answering every actor separately, a firm can get good at a handful
of themes --- and the datapoints beneath them --- and thereby respond to most
of its demands at once. The analysis that follows shows precisely which themes
those are and which datapoints they require.


%LI: good - only problem is - should I make business case for sustainability more? why it is important to build this capability - find more sources of who asks and stories of suppliers being cut because cannot answer - etc? or good as is?