% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
]{article}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{longtable,booktabs,array}
\newcounter{none} % for unnumbered tables
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\ifLuaTeX
  \usepackage{luacolor}
  \usepackage[soul]{lua-ul}
\else
  \usepackage{soul}
\fi
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{}

\begin{document}

\begin{titlepage}
  \centering
  \vspace*{4cm}

  {\LARGE\textbf{Navigating the Sustainability Disclosure Ecosystem}\par}
  \vspace{1cm}
  {\large\emph{A diagnostic framework for European seafood processing SMEs}\par}
  \vspace{0.5cm}
  {\large\emph{Moving from reactive compliance to proactive sustainability strategy}\par}
  \vspace{2cm}
  {\large\textbf{Introduction and Methodology --- Working Paper}\par}
  \vspace{2cm}
  {\large Linne Verhoeven\par}
  \vspace{0.5cm}
  {\large May 2026\par}
\end{titlepage}

\section{1. Introduction}\label{introduction}

\subsection{1.1 Sustainability as
strategy}\label{sustainability-as-strategy}

Companies are increasingly faced with sustainability demands originating
from regulatory, market, and societal stakeholders. The way firms
respond to these demands varies along a continuum of sustainability
maturity, between two ideal-typical postures (Russo \& Fouts, 1997;
Sharma \& Vredenburg, 1998).

A reactive approach implies that the firm meets the bare minimum of
stakeholder demands by acting only when external pressure arises.
Sustainable practices are not specifically developed and remain
peripheral to the business model. Firms with a reactive posture comply
with the minimum sustainability prerequisites required to operate but
are unlikely to benefit from performance gains associated with deeper
sustainability integration. A proactive approach, in contrast, implies
voluntary engagement that exceeds the legal minimum, contributing to
environmental and social development as well as long-term economic
performance. Empirical evidence from CSR research in small and
medium-sized enterprises (SMEs) suggests that proactive economic-related
CSR generates positive direct effects on financial performance over the
long term (Torugsa et al., 2013).

The barrier to proactive engagement, particularly for
resource-constrained firms, is not lack of intent but external-input
overload. Fragmented and complex sustainability governance pushes SMEs
toward reactive compliance behaviour rather than strategic integration
(Chan et al., 2022). Where multiple stakeholders demand different
information through different channels, the practical effect on SMEs
with limited compliance capacity is paralysis or minimum-viable
response.

\subsection{1.2 SME sustainability
constraints}\label{sme-sustainability-constraints}

SMEs face a well-documented set of constraints that shape their
sustainability strategy choices. These include: limited financial and
managerial resources; lack of dedicated internal expertise in
environmental and social reporting; weak data systems and IT
infrastructure; high dependence on a small number of B2B customers; and
uncertainty regarding the evolving regulatory landscape (Battisti \&
Perry, 2011).

In practice these constraints mean that an SME confronted by multiple,
overlapping sustainability requirements cannot simply expand its
compliance function proportionally. Each new disclosure request from a
customer, certifier, investor, or regulator imposes a fixed overhead
disproportionate to the firm\textquotesingle s size. The resulting
incentive is to treat sustainability disclosure as transactional
compliance, answering questionnaires as they arrive, rather than as a
strategic investment in capabilities that could serve multiple demands
simultaneously (Interview).

\subsection{1.3 Fragmented governance in the seafood
sector}\label{fragmented-governance-in-the-seafood-sector}

Seafood is the most highly traded food product globally, with per capita
consumption increasing more rapidly than other animal proteins (Barklay
and Miller., 2018). The sector is structurally complex, involving global
sourcing networks, multiple processing stages, and exposure to a wide
range of sustainability risks including greenhouse gas emissions, marine
ecosystem impacts, antimicrobial use, forced labour in fishing fleets,
and food safety incidents (Barklay and Miller., 2018).

European seafood processors face this complexity from a position of
compounded pressure. In addition to horizontal European legislation
addressing waste and packaging (Regulation (EU) 2025/40), sustainability
claims (the proposed Empowering Consumers Directive, Directive (EU)
2024/825), and corporate sustainability disclosure (CSRD/ESRS, CSDDD),
the sector is subject to a dense overlay of sector-specific regulation:
the EU Common Fisheries Policy (Regulation (EU) No~1380/2013), the IUU
Regulation (Council Regulation 1005/2008), the recast Control Regulation
(2023/2842), the EU Deforestation Regulation (2023/1115, relevant for
aquaculture feed sourcing), and the Farm to Fork Strategy under the
European Green Deal.

Layered on top of this regulatory architecture is a fragmented landscape
of voluntary instruments: certifications (MSC for wild-capture, ASC and
BAP for aquaculture, GLOBALG.A.P.), food safety standards (IFS Food,
BRCGS, FSSC 22000, benchmarked by GFSI), supply-chain transparency
platforms (Sedex/SMETA, EcoVadis), investor frameworks (CDP, SBTi, TCFD,
IFRS S2, the Coller FAIRR Protein Producer Index), industry initiatives
(GSSI, GDST, Move to Minus 15, ISSF for tuna), and management-system
standards (ISO 14001, 45001, 50001). Governance is structurally
fragmented, with no single coordinating actor and overlapping but
non-identical demands from each (Bush \& Oosterveer, 2019).

The empirical implication, surfaced in early scoping interviews for this
research, is that European seafood SMEs receive sustainability data
requests from multiple actors who individually demand reasonable
information but collectively impose an unsustainable response burden.
The SME response, predictably, is reactive: comply with the loudest
demand, defer the rest, do not invest in capability that would serve
more demands than the one in front of them.

\subsection{1.4 The sustainability reporting
landscape}\label{the-sustainability-reporting-landscape}

The European Union has positioned corporate sustainability reporting as
a key lever for the European Green Deal, intending to improve the
availability, comparability, and reliability of sustainability
information available to investors, regulators, and value-chain partners
(European Commission, 2019). Disclosure is conceptualised not only as a
transparency mechanism but as a tool that shapes economic behaviour by
directing capital toward sustainable activities.

The Corporate Sustainability Reporting Directive (CSRD; Directive (EU)
2022/2464) operationalises this ambition. CSRD is implemented through
the European Sustainability Reporting Standards (ESRS; Commission
Delegated Regulation (EU) 2023/2772), which define mandatory disclosure
requirements across twelve cross-cutting and topical standards covering
environmental, social, and governance dimensions. ESRS introduces the
double materiality principle, requiring undertakings to report both on
how sustainability matters affect their financial performance and on the
firm\textquotesingle s own impacts on the economy, environment, and
people (EFRAG, 2021).

CSRD scope is restricted to large undertakings and listed entities
meeting specified employee, turnover, and balance-sheet thresholds.
Under the Omnibus simplification proposals (European Commission, 2025),
the scope of CSRD has been further restricted, with smaller
previously-in-scope firms now exempted. The majority of European seafood
processing companies fall below CSRD thresholds and are therefore not
directly subject to ESRS reporting obligations.

To address the disclosure needs of out-of-scope SMEs, EFRAG has
developed the Voluntary Sustainability Reporting Standard for SMEs
(VSME), comprising a basic and a comprehensive module. VSME is designed
to be proportionate to SME constraints, reducing the disclosure load
from approximately 1,000 ESRS datapoints to approximately 100. Adoption
is voluntary, but the framework is intended to provide a recognised
reference set that satisfies the most common information requests SMEs
receive from value-chain partners and financial institutions.

In practice, however, two structural gaps are observable. First, the
leap from no sustainability reporting at all to even the basic VSME
module remains significant for many SMEs (interview). Second, the gap
between VSME and full ESRS is substantially larger and is precisely the
territory in which non-CSRD seafood SMEs operate when responding to
ad-hoc customer questionnaires, certification audits, and investor
requests. The sustainability reporting frame, in other words, does not
cover the operational reality of European seafood SMEs in the post-CSRD
environment.

\subsection{1.5 Research gap and the diagnostic-tool
argument}\label{research-gap-and-the-diagnostic-tool-argument}

Existing literature addresses several aspects of this problem in
isolation. Sustainability reporting maturity models describe staged
organizational evolution in reporting practice (Baumgartner et al.,
2010; Cöster et al., 2020). Studies of CSRD implementation costs
document the burden faced by in-scope firms (EFRAG, 2025).
Stakeholder-pressure literature explains why firms respond to particular
demands (Mitchell, Agle \& Wood, 1997; Suchman, 1995).
Institutional-isomorphism theory provides a lens on why disclosure
conventions converge despite formal voluntariness (DiMaggio \& Powell,
1983). The seafood-specific literature documents the fragmented
private-governance landscape (Bush et al., 2017; Tlusty, 2012;
Vandergeest, 2018).

What is missing is an integrated empirical analysis that (i) maps the
actual cross-framework disclosure demand facing European seafood SMEs,
(ii) identifies where multiple stakeholders converge versus where they
diverge, and (iii) translates that empirical pattern into a diagnostic
tool that lets an SME self-position on the reactive-to-proactive
spectrum and identify which capability investments would most
efficiently expand its sustainability response.

This thesis addresses that gap. The empirical material is drawn from the
disclosure practice of five large European seafood processors that are
CSRD-reporting; this cohort defines the realistic upper bound of
seafood-sector disclosure depth. Cross-referenced against the broader
stakeholder ecosystem of regulators, customers, investors, and industry
initiatives, the analysis identifies disclosure topics that face
concentrated cross-stakeholder demand. These hotspots define the
high-leverage capability investments for SMEs: topics where building
reporting capability once satisfies many demands simultaneously,
enabling the SME to shift from reactive question-by-question response to
proactive capability-driven engagement.

\section{2. Research questions and
contribution}\label{research-questions-and-contribution}

\subsection{2.1 Research questions}\label{research-questions}

The research is organised around three interconnected questions:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{RQ1 - Empirical:} What ESRS topical disclosure requirements
  are actually reported on by mature European seafood processing
  companies, and what does this empirical cohort reveal about the
  practical upper bound of sector disclosure depth?
\item
  \textbf{RQ2 - Analytical:} Across the full landscape of regulatory,
  customer, investor, and industry frameworks active in the European
  seafood sector, which disclosure topics constitute hotspots of
  concentrated cross-stakeholder demand?
\item
  \textbf{RQ3 - Prescriptive:} How can a European mid-sized seafood
  processor diagnose its current sustainability posture (reactive vs
  proactive) and prioritize capability investments to progress along
  that spectrum in a way that satisfies the most pressing
  cross-stakeholder demands?
\end{enumerate}

\subsection{2.2 Intended contributions}\label{intended-contributions}

The thesis contributes to three audiences:

\begin{itemize}
\item
  \textbf{For seafood SMEs (primary audience):} a diagnostic framework
  and a stage-of-development roadmap that translates the complexity of
  the disclosure ecosystem into a sequence of capability investments
  tailored to the SME\textquotesingle s current posture and identified
  stakeholder pressures.
\item
  \textbf{For academic literature:} a first systematic mapping of
  cross-stakeholder disclosure demand at the sector level using a
  defined coverage scale. An extension of sustainability-reporting
  maturity-model literature with a diagnostic-tool operationalization.
  An empirically grounded application of the reactive/proactive CSR
  distinction.
\item
  \textbf{For EU policy (secondary):} empirical input to the ongoing
  simplification debate (Omnibus, VSME extension, sector standards
  postponement). The cross-stakeholder hotspot analysis identifies where
  the regulatory and voluntary disclosure ecosystem is converging and
  therefore where coordination or capping of demand would deliver
  disproportionate burden reduction for SMEs. This aligns with the
  Competitiveness Compass target of cutting administrative burden by
  $\geq$35\% for SMEs (European Commission, 2025).
\end{itemize}

\section{3. Conceptual framework}\label{conceptual-framework}

\subsection{3.1 The reactive-proactive sustainability
spectrum}\label{the-reactive-proactive-sustainability-spectrum}

The reactive-proactive distinction (Sharma \& Vredenburg, 1998; Torugsa
et al., 2013) provides the central organising frame. The two postures
differ along several dimensions:

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.2564}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3718}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3718}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{Dimension}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Reactive posture}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Proactive posture}
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Trigger for action} & External demand (questionnaire, audit,
regulatory deadline) & Strategic anticipation; voluntary commitments \\
\textbf{Disclosure approach} & Transactional, question-by-question;
minimum viable & Capability-driven, integrated; goes beyond legal
minimum \\
\textbf{Data infrastructure} & Ad-hoc, manual, rebuilt per request &
Systematic, automated, reusable across demands \\
\textbf{Strategic positioning} & Compliance cost center & Competitive
differentiation; risk-adjusted financial benefit \\
\textbf{Stakeholder relationship} & Defensive; minimum communication &
Proactive engagement; trust-building \\
\end{longtable}
}

(for now made by me based on: Sharma \& Vredenburg, 1998; Torugsa et
al., 2013)

Empirical evidence in CSR research indicates that the proactive posture
is associated with positive long-term financial performance effects
(Torugsa et al., 2013). The barrier to adoption among
resource-constrained SMEs is not rejection of the proactive posture in
principle, but the absence of a tractable pathway from reactive practice
to proactive capability given the fragmented and unpredictable nature of
external demand (interview -- look for source).

\subsection{3.2 Stakeholder pressure and fragmented
governance}\label{stakeholder-pressure-and-fragmented-governance}

Stakeholder-salience theory (Mitchell, Agle \& Wood, 1997) explains how
firms triage competing claims based on the legitimacy, urgency, and
power of each stakeholder. In contexts of single-stakeholder dominance,
the response is clear. In contexts of multiple legitimate stakeholders
with overlapping but non-identical demands, case in point for the
seafood sustainability ecosystem, the absence of demand coordination
creates the fragmentation problem: each stakeholder is individually
legitimate, but the cumulative response burden is unmanageable.

Institutional theory (DiMaggio \& Powell, 1983) further explains why
disclosure conventions converge across firms even where formally
voluntary: coercive isomorphism (regulatory pressure), mimetic
isomorphism (copying perceived leaders), and normative isomorphism
(professionalisation of sustainability functions) all push firms toward
similar disclosure choices. For SMEs the mimetic mechanism is
particularly relevant because they typically lack the in-house expertise
to develop independent positions.

The fragmented governance of the seafood sustainability movement, with
regulator, brands, retailers, certifiers, NGOs, and investors all
exercising parallel and overlapping demands (Bush et al., 2019),
represents an extreme case of the multi-stakeholder challenge. This
thesis takes as its starting analytical premise that the SME response to
such fragmentation is structurally reactive, and that this reactivity is
not a failure of SME intent but a rational adaptation to incoherent
external demand.

\subsection{3.3 The bridge: cross-stakeholder demand
convergence}\label{the-bridge-cross-stakeholder-demand-convergence}

The conceptual bridge from reactive to proactive in this fragmented
environment lies in identifying disclosure topics where multiple
stakeholders converge. Where the same topic is demanded by a customer
(e.g., via a chain-of-custody certification), an investor (via a CDP
questionnaire), an industry framework (via GRI 13), and a regulator (via
ESRS or a sector-specific instrument), the SME that invests in
capability for that topic obtains compound returns: a single capability
investment satisfies multiple demands.

Hotspots of cross-stakeholder convergence therefore represent the
strategic-investment opportunities through which an SME can begin to
move from reactive to proactive. Building disclosure capability on a
hotspot topic delivers, by definition, leverage: each unit of capability
invested generates response capacity across multiple stakeholders. By
contrast, building capability on a topic demanded by only one
stakeholder type generates narrower returns and entrenches the
transactional posture.

This insight motivates the central analytical construct of the thesis: a
heatmap of disclosure topics by stakeholder type, populated from a
systematic mapping of the seafood sustainability ecosystem. The hotspots
identified in this heatmap form the empirical basis for a
stages-of-development roadmap and a diagnostic tool through which an SME
can position itself, identify its highest-leverage next investment, and
progressively expand its sustainability response.

\subsection{3.4 The seafood sustainability playing field: an integrative
framework}\label{the-seafood-sustainability-playing-field-an-integrative-framework}

Sections 3.1 to 3.3 establish three building blocks: the
reactive-proactive continuum along which firms position themselves
(3.1); the fragmented, multi-stakeholder structure of sustainability
pressure in the seafood sector (3.2); and the principle that
cross-stakeholder convergence creates leverage (3.3). This section
integrates these building blocks into a single organising framework, the
seafood sustainability playing field, which constitutes the central
conceptual contribution of this thesis. The framework serves two
purposes: it provides a structured map of the disclosure environment in
which seafood firms operate, and, through the construct of capability
corridors, it identifies where compliance with mandatory requirements
can be deliberately extended into proactive sustainability strategy.

\subsubsection{The three structuring
axes}\label{the-three-structuring-axes}

The playing field organises the disclosure environment along three axes.
The first two are structural, in that they define the position of every
requirement, while the third is an attributional overlay.

The necessity axis is the framework's spine. It positions each
requirement on a spectrum of obligation that runs from a foundational
floor up through four tiers. Beneath the four tiers sits Tier 0, food
integrity: the non-negotiable food-safety and food-law floor - General
Food Law, the hygiene regulations, and the customer-required food-safety
certifications IFS Food, BRCGS and FSSC 22000 - that every EU seafood
processor clears before any sustainability question arises. Tier 0 is in
scope of the firm but is held in a separate band rather than treated as
a sustainability requirement, because it is a precondition of operating.
Above it, Tier 1, legal license to operate, comprises requirements
without which the firm cannot lawfully operate, such as the
catch-certificate regime of the EU IUU Regulation, the Fisheries Control
Regulation, the EU Deforestation Regulation and packaging law; the
Common Fisheries Policy frames this regime but binds Member States
rather than the processor directly. Tier 2, market license to operate,
comprises requirements that are legally optional but commercially
non-negotiable, because no major retail customer will list a supplier
without them, such as MSC or ASC chain-of-custody certification for
particular products and the SMETA and EcoVadis social-audit memberships.
Tier 3, competitive expectation, comprises requirements that are
increasingly normalised, such that their absence constitutes a
competitive disadvantage, such as VSME reporting and the value-chain
disclosures cascaded from customers under CSRD and its ESRS standards.
Tier 4, strategic differentiation, comprises genuinely voluntary
commitments that signal leadership. Alongside the tiers, a separate
enabling layer (Layer 2) holds the methods, management systems and data
infrastructure - the GHG Protocol, the ISO management-system standards,
life-cycle-assessment methods - whose necessity is derived from the
demand they serve rather than being intrinsic, and which therefore sit
outside the necessity grid. This axis is grounded in legitimacy theory
(Suchman, 1995), whose distinction between pragmatic, moral and
cognitive legitimacy corresponds to the progression from legally
enforced to normatively expected obligation, and in institutional theory
(DiMaggio \& Powell, 1983), whose coercive, mimetic and normative
isomorphism mechanisms correspond to the lower, middle and upper tiers
respectively. The necessity axis is structurally central because it maps
directly onto the reactive-proactive continuum of Section 3.1: a
reactive firm operates only within the foundation and Tiers 1 and 2,
whereas a proactive firm extends its engagement into Tiers 3 and 4.

The domain axis classifies each sustainability requirement by its
subject matter, distinguishing two categories: seafood-sector
sustainability (the sector-specific environmental and social topics such
as bycatch, stock health, aquaculture welfare, and chain-of-custody
traceability) and general corporate sustainability (the cross-sector
topics such as greenhouse-gas emissions, energy, labour, and
governance). Core operational and food-integrity requirements - food
safety, hygiene, basic legal compliance - are not treated as a third
domain but as the Tier 0 foundation described above, because they are a
precondition of operating rather than a position on the sustainability
spectrum. The domain axis thus presents the distinction between
seafood-specific and general sustainability requirements as two adjacent
columns, while food integrity underlies both.

The source axis attributes each requirement to its primary stakeholder
driver: regulator, customer, investor, or industry and civil society.
Unlike the first two axes, source is not used to structure the framework
but is applied as a colour-coded overlay, because stakeholder drivers
cross-cut both necessity and domain. Source is assigned by a single
test: does the instrument legally bind the processor? If it does, the
driver is the regulator (for example the EU Deforestation Regulation,
packaging law, or occupational-safety law). If it does not, the driver
is the stakeholder that transmits the demand - a customer audit codes to
customer, a capital-market expectation to investor - and self-adopted
voluntary frameworks are coded by their primary intended audience, with
investor-facing standards such as IFRS S1/S2 and TNFD coding to investor
and multistakeholder or NGO-led standards such as GRI and SBTi coding to
industry and civil society. This axis draws on stakeholder-salience
theory (Mitchell, Agle \& Wood, 1997): the same requirement may be
demanded with differing power, legitimacy and urgency by different
actors.

A second overlay, obligation locus, records separately from source
whether the processor is itself the bound party. A requirement is direct
where the processor is the obligated, certified or reporting entity;
indirect where the obligation sits elsewhere in the value chain and
reaches the processor through its own disclosure duty or a buyer's
demand, in which case it is placed at the tier of that transmission
mechanism; and context where the instrument binds a Member State or
another actor with no mechanism that transmits an obligation to the
processor at all. Source and locus are orthogonal. The EU Deforestation
Regulation is regulator-driven and direct; the fisheries-labour
Directive is customer-driven, because it binds vessel owners and reaches
the processor through social audits, yet its locus is indirect; the
voluntary VSME standard is investor-driven and direct; and CSRD reaches
the sub-threshold processor as a customer-driven, indirect value-chain
demand. Holding the two dimensions apart prevents the common error of
reading the legally bound party off the identity of the actor who
actually applies the pressure.

Rendering source and locus as overlays rather than axes makes visible a
pattern central to the argument of this thesis, namely that necessity,
domain, source and locus are correlated but not identical. Tier 1
requirements are predominantly regulator-driven and direct, Tier 2
predominantly customer-driven, and the voluntary upper tiers spread
across investor and industry drivers. It is the visible gradient of this
correlation, rather than its assumption, that the framework contributes.
Figure 1 presents the populated framework, with the necessity tiers as
rows, the two domain categories as columns, each instrument positioned
in its cell and colour-coded by source, and its obligation locus marked.

\subsubsection{Capability corridors and the compliance-to-strategy
synergy}\label{capability-corridors-and-the-compliance-to-strategy-synergy}

A grid of requirements organised by necessity and domain is a
descriptive map; it does not, by itself, explain how a firm moves from
reactive compliance to proactive strategy. The construct that performs
this function is the capability corridor. A capability corridor is an
underlying organisational capability, a body of data, a process, or an
information system, upon which multiple requirements across different
cells of the framework draw. The construct is grounded in the
resource-based view of the firm (Barney, 1991) and the
dynamic-capabilities literature (Teece, 2007): a capability, once built,
is fungible across the demands that require it.

Two corridors illustrate the construct. The traceability corridor runs
from a Tier 1 anchor, the catch-certificate and lot-tracking obligations
of the IUU and Control Regulations, through Tier 2 chain-of-custody
certification (MSC, ASC) and the GDST traceability standard, into Tier 3
value-chain disclosure under VSME and ESRS. The supplier-engagement
corridor runs from a Tier 2 anchor, the supplier-control requirements of
food safety certifications such as IFS Food and BRCGS, through SMETA
social auditing into Tier 3 value-chain worker disclosure. In each case
a single capability, built to satisfy a lower-tier mandatory
requirement, also supplies a substantial share of the infrastructure
required by higher-tier strategic disclosures.

This observation yields the framework's central strategic proposition.
Because a capability built for a lower-tier mandatory requirement
subsidises the higher-tier requirements that draw on the same corridor,
the marginal cost of proactive sustainability disclosure is considerably
lower than it appears when each requirement is assessed in isolation. A
firm that builds traceability capability because Tier 1 law compels it
has, in doing so, already financed a substantial portion of the
infrastructure required for MSC certification and for ESRS value-chain
disclosure. The proactive sustainability strategy, on this account, is
to a large degree the deliberate recognition and extension of
capabilities the firm was compelled to build for legal survival. The
reactive firm, by contrast, rebuilds capability requirement by
requirement, never recognising the corridor, and so perpetually pays the
full isolated cost of each demand. The corridors are therefore where
compliance becomes strategy, and they constitute the conceptual bridge,
anticipated in Section 3.3, between the reactive and proactive postures.

\subsubsection{Relationship to the empirical
analysis}\label{relationship-to-the-empirical-analysis}

The playing field is a conceptual framework; the empirical chapters of
this thesis populate and test it. Phases 1 and 2 of the methodology
(Sections 4.3 and 4.4) identify the disclosure topics and the
instruments that occupy the framework's cells. Phase 3 (Section 4.5)
codes the demand each instrument places on each topic, producing the
cross-framework demand matrix that empirically substantiates the source
overlay and reveals where corridors run. Phase 4 (Section 4.6) operates
on this matrix to identify cross-stakeholder hotspots and to assign
topics to development stages.

A clarification is required regarding the relationship between the
framework's necessity tiers and the development stages produced by the
Phase 4 analysis, because the two are related but distinct. The
necessity tiers are a structural property of each requirement, namely
how obligatory it is, and form the rows of the framework. The
development stages are an empirical product of the demand analysis,
namely how many stakeholder types converge on a topic, and identify
where leverage is concentrated. A topic may be high in necessity but
narrow in cross-stakeholder demand, or low in necessity but broad in
demand. The roadmap that this thesis ultimately proposes therefore
sequences capability investment by necessity first and by
cross-stakeholder leverage second: a firm addresses its Tier 1
obligations as a precondition of operating, and, among the topics within
reach at each tier, prioritises those that sit on high-leverage
corridors. Necessity defines what must be done; the demand analysis
defines, among what must or could be done, where to begin.

\section{4. Methodology}\label{methodology}

\subsection{4.1 Research design
overview}\label{research-design-overview}

The research employs a mixed-methods empirical design organised in five
phases:

\begin{itemize}
\item
  \textbf{Phase 1 - Universe construction.} Defining the set of
  disclosure topics in scope for the analysis, drawn from cohort
  empirical practice (ESRS-derived) and supplemented with
  sector-material topics outside ESRS (GRI).
\item
  \textbf{Phase 2 - Landscape mapping.} Cataloguing the regulatory,
  customer-driven, investor-driven, and industry-driven frameworks
  active in the European seafood sustainability ecosystem.
\item
  \textbf{Phase 3 - Cross-framework demand mapping.} Coding each
  disclosure topic against each priority framework on a defined coverage
  scale, producing the central demand matrix.
\item
  \textbf{Phase 4 - Hotspot analysis and stage assignment.} Identifying
  topics with concentrated cross-stakeholder demand. Assigning topics to
  development stages; constructing the diagnostic tool.
\item
  \textbf{Phase 5 - Case study validation.} Applying the framework to
  one or two SME case studies to test usability and refine the
  diagnostic tool.
\end{itemize}

\subsection{4.2 Empirical scope: the large-reporter
cohort}\label{empirical-scope-the-large-reporter-cohort}

The empirical foundation is the disclosure practice of five large
European seafood processing companies that are CSRD-reporting or have
published sustainability reports of comparable depth: Bolton Group, A.
Espersen A/S, Mowi ASA, Nomad Foods Limited, and Profand Group. These
five represent the realistic upper bound of European seafood sector
disclosure depth and span both wild-capture and aquaculture operations,
branded and private-label, listed and unlisted.

The data sources are each company\textquotesingle s most recent publicly
available sustainability report (2023 or 2024 reporting cycle),
supplemented where available by their CSRD-aligned ESRS-index
appendices. The unit of empirical analysis is the ESRS Disclosure
Requirement (DR); each report was coded for whether the company reports
against each of the 70 ESRS topical DRs (E1--E5, S1--S4, G1).

\subsection{4.3 Phase 1 - Universe
construction}\label{phase-1---universe-construction}

The disclosure-topic universe used in the cross-framework mapping is
constructed in four steps.

\textbf{Step 1.1 Cohort coverage matrix}

A binary coverage matrix is built recording, for each ESRS topical DR
and each of the six cohort companies, whether the DR is materially
reported against. For the three GRI-reporting companies, disclosures are
first translated onto the ESRS DR scale through the GRI--ESRS bridge, so
the matrix is lens-unified across all six companies. This matrix is the
empirical record of disclosure practice in the cohort.

\textbf{Step 1.2 Consensus filter: the 26-DR core and topic-level
resolution}

The core universe is built from the bridged coverage matrix by a
unit-aware consensus rule applied in two tiers, both operating at the
same $\geq$5-of-6 evidentiary threshold.

Tier 1 --- DR-level consensus. A Disclosure Requirement is admitted to
the universe if reported by at least five of the six cohort companies.
Reporting by $\geq$5/6 is treated as strong empirical evidence that the DR is
both material to the sector and feasible to report on at scale. Applied
to the bridged coverage matrix, this rule yields a consensus core of 26
DRs.

Tier 2 --- Topic-level resolution. Because the GRI--ESRS bridge places
all six companies on a common DR scale, reporting that is dispersed
across several DRs within one topical area can fall below the DR-level
threshold even when most companies report the topic. To correct this
granularity artefact, the firm-union is computed at topic level for each
ESRS topical standard: a topic is admitted as a single topic-level row
if at least five of the six companies report on at least one DR within
that topic. Applied uniformly across all topical standards, this rule
admits two additional topics:

$\bullet$ Biodiversity (ESRS E4) --- 5 of 6 companies report at least one of
E4-1 through E4-5, with reporting dispersed across the policy, action,
target and metric DRs such that no individual sub-DR clears the $\geq$5/6
bar.

$\bullet$ Consumer health and food safety (ESRS S4) --- 5 of 6 companies report
at least one S4 DR, again dispersed across sub-DRs.

GRI 13 (Agriculture, Aquaculture and Fishing Sectors 2022) corroborates
the sector-materiality of both topics. Because these additions enter at
the same evidentiary standard as Tier 1, no relaxed or ad-hoc rescue
threshold is introduced. The high-confidence universe at the close of
Step 1.2 contains 28 entries: the 26 Tier 1 DRs plus the two Tier 2
topics.

A consequence of the two-tier rule is that 26 entries are coded at DR
level and 2 at topic level in the cross-framework demand matrix of Phase
3. This asymmetry is treated as a feature rather than a flaw: the
granularity of each entry's row in the demand matrix matches the
granularity at which cohort evidence supports its inclusion. A
Coding\_Level column is carried alongside each row to make the
granularity transparent throughout the analysis.

\textbf{Step 1.3 GRI 13 sector cross-check}

The 28-entry high-confidence universe is cross-referenced against the 26
likely-material topics enumerated in GRI 13: Agriculture, Aquaculture
and Fishing Sectors 2022. Each GRI 13 topic is classified into one of
three categories:

$\bullet$ Covered --- at least one mapped ESRS DR is present in the
high-confidence universe.

$\bullet$ Reporting gap --- a mapped ESRS DR exists but is not in the universe,
indicating that the cohort under-reports the topic relative to its
sector materiality per GRI 13.

$\bullet$ Structural gap --- no ESRS DR exists for the topic at all, indicating
that ESRS itself does not cover what GRI 13 considers sector-material.

The reporting-gap and structural-gap categories together describe the
topics for which an ESRS-centric universe would understate the
disclosure environment that European seafood SMEs actually face.

\textbf{Step 1.4 Sector-material topics carried into the demand matrix}

The cross-check in Step 1.3 is not left as commentary. The five topics
that are most sector-defining for seafood yet have no clean ESRS (and,
equally, no VSME) home are carried into Phase 3 as a separate,
explicitly flagged block of sector-material rows: supply-chain
traceability (chain of custody), wild-stock status and sustainable wild
sourcing, feed and marine-ingredient sourcing, antimicrobial use, and
animal welfare. Wild-stock status partly overlaps ESRS E4 and is
retained as a distinct sourcing-decision row; the remaining four have no
dedicated Disclosure Requirement in either ESRS or the VSME standard.

These five are deliberately not derived from cohort consensus --- the
cohort does not report them at scale precisely because the reporting
taxonomy does not ask for them. They are admitted on a different,
clearly labelled basis: sector materiality per GRI 13 and demand by EU
sector legislation or major non-regulatory frameworks active in the
seafood ecosystem (e.g. the IUU and Fisheries Control Regulations,
MSC/ASC chain of custody, GDST, FisheryProgress). Carrying them as a
flagged block keeps the cohort-evidenced core and the sector-material
additions epistemically distinct, while ensuring the demand matrix
reflects the disclosure environment a seafood SME actually faces rather
than an ESRS-only abstraction. In the matrix these rows score near-zero
across the ESRS/VSME and investor columns and concentrate on the seafood
certification and regulatory columns --- the contrast is itself a
finding: an SME preparing only for VSME would miss the topics its buyers
and regulators most demand.

\textbf{Final universe}

The disclosure-topic universe used as the row dimension of the
cross-framework demand matrix in Phase 3 comprises 33 rows:

$\bullet$ 26 Tier 1 entries --- DR-level cohort consensus ($\geq$5/6);

$\bullet$ 2 Tier 2 entries --- topic-level cohort consensus ($\geq$5/6): E4
Biodiversity and S4 Consumer health and food safety;

$\bullet$ 5 sector-material entries --- structural-gap seafood topics carried in
on GRI 13 / sector-legislation grounds, flagged as additions rather than
cohort-evidenced.

Each entry is carried into Phase 3 with metadata indicating its
inclusion tier, its coding level (DR or Topic), and a confidence flag
distinguishing the 28 cohort-evidenced high-confidence entries from the
5 sector-material additions.

\subsection{4.4 Phase 2 - Landscape
mapping}\label{phase-2---landscape-mapping}

Phase 2 catalogues the full landscape of frameworks, standards,
certifications and initiatives that place sustainability-disclosure
demands on a European seafood-processing SME, and arranges them into the
necessity-by-domain map that populates the playing field of Section 3.4.
The cohort sustainability reports are the empirical entry point: every
framework explicitly referenced by at least one cohort company is
included, and desk research enriches the metadata and adds instruments
that are binding on the entity but absent from cohort disclosure.
Necessity is assessed throughout from the perspective of the SME - a B2B
supplier of 50 to 250 employees selling to large EU retailers - so the
same instrument may sit at a different tier for a differently situated
firm.

\textbf{The necessity tiers and domains.}

\hl{\emph{Key Question: What happens to the firm without it?}
Can\textquotesingle t lawfully operate $\to$ Tier 1. Customers
won\textquotesingle t list you $\to$ Tier 2. Falling behind peers $\to$ Tier 3.
Pure leadership signal $\to$ Tier 4.}

Each instrument is placed in one cell of a grid whose rows are necessity
tiers and whose columns are the two sustainability domains. The rows are
a Tier 0 foundation of food-integrity law and food-safety certification,
which is in scope of the firm but excluded from the disclosure analysis
and held in a separate band; Tier 1, legal license to operate, of
instruments that legally bind the processor; Tier 2, market license, of
requirements that are legally optional but commercially non-negotiable;
Tier 3, competitive expectation; and Tier 4, strategic differentiation.
The columns are seafood-sector sustainability and general corporate
sustainability. A separate enabling layer, Layer 2, holds the
measurement methods, management systems and data infrastructure - for
example the GHG Protocol, the ISO 14001/45001/50001 standards and
life-cycle-assessment methods - whose necessity is derived from the
demand they serve and which therefore sit outside the grid.

\textbf{Source: the binds-the-processor test.}

\hl{\emph{Key Question: Who drives the obligation to the processor?}
Decision rule: Does the instrument legally bind the processor?
\textbf{Yes $\to$ Regulator.} \textbf{No $\to$ the stakeholder that transmits
the demand} (buyer audit/questionnaire $\to$ Customer; capital-market
expectation $\to$ Investor; self-adopted voluntary framework $\to$ by intended
audience)}

Each instrument is assigned a single primary stakeholder driver by a
pinned test, so that the coding is reproducible rather than
impressionistic. The test asks whether the instrument legally binds the
processor; if it does, the source is the regulator, and if it does not,
the source is the stakeholder that actually transmits the demand to the
firm. The four codes are:

\begin{itemize}
\item
  \textbf{Regulator:} the instrument legally binds the processor - for
  example the IUU and Fisheries Control Regulations, the EU
  Deforestation Regulation, the Packaging and Packaging Waste
  Regulation, and occupational-safety and working-time law.
\item
  \textbf{Customer:} the instrument does not bind the processor but is
  transmitted by retail or B2B buyers through audits, supplier
  questionnaires and value-chain data requests - for example MSC, ASC,
  BAP, GLOBALG.A.P., Sedex/SMETA and EcoVadis, and, for a sub-threshold
  SME, the CSRD/ESRS, CSDDD and CDP demands cascaded from large
  customers.
\item
  \textbf{Investor:} the instrument is driven by capital-market
  expectations and its primary intended user is a provider of capital -
  SFDR, the EU Taxonomy, IFRS S1/S2, the FAIRR and Dow Jones indices,
  and VSME, which a sub-threshold SME adopts largely at the prompting of
  lenders and banks.
\item
  \textbf{Industry / NGO:} self-adopted voluntary frameworks whose
  primary audience is multistakeholder or NGO-led - GRI (incl. GRI 13),
  SBTi, the UN Global Compact and SDGs, GDST, FIPs/FisheryProgress, ISSF
  and SeaBOS.
\end{itemize}

\textbf{Obligation locus.}

\hl{\emph{Key Question: Is the processor itself the bound party?}
Decision rule: Processor is the obligated/certified/reporting entity $\to$
\textbf{Direct ($\bullet$)}. Obligation sits elsewhere in the chain and reaches
the processor via a buyer demand or its own disclosure duty $\to$
\textbf{Indirect ($\circ$)}. Binds a Member State or other actor with no
transmission to the processor $\to$ \textbf{Context (---)}}

Orthogonally to source, each instrument carries an obligation-locus code
that records whether the processor is itself the bound party, encoded in
the map as a filled marker for direct, an open marker for indirect, and
a dash for context. A requirement is direct where the processor is the
obligated, certified or reporting entity - the EUDR due-diligence duty,
MSC or ASC certification of the seafood it sells, its own VSME report.
It is indirect where the obligation sits elsewhere in the chain and
reaches the processor through a buyer demand or its own disclosure duty,
in which case it is placed at the tier of that transmission mechanism -
the fisheries-labour Directive, which binds vessel owners, or CSRD and
CSDDD, which bind large undertakings and cascade down. It is context
where the instrument binds a Member State or another actor with no
mechanism transmitting an obligation to the processor - the Common
Fisheries Policy, the Marine Strategy Framework Directive, RFMO catch
rules, the Water and Waste Framework Directives - and these are retained
in-cell as greyed context rather than active demands. Source and locus
are independent: an instrument can be customer-driven yet indirect (the
fisheries-labour Directive) or regulator-driven yet context (the Common
Fisheries Policy).

\textbf{Scoping and consolidation decisions.}

Four coding decisions are applied consistently and recorded with the
map. First, dual-channel instruments that reach a sub-threshold SME
through both customers and capital markets - CSRD/ESRS, CSDDD and CDP -
are coded Customer, because for a firm below the CSRD threshold the
dominant transmitter is the large customer's value-chain data and
due-diligence cascade rather than its own investors; the Investor code
is reserved for instruments whose primary intended user is capital.
Second, certification locus follows a core-product test: a scheme is
direct only where the processor holds it for the seafood it sells (MSC,
ASC, BAP, GLOBALG.A.P.) and indirect where the certified material is a
non-core input the processor does not physically handle. Third, on that
logic RSPO is demoted from a first-class Tier 2 entry to a conditional
footnote, because palm reaches a seafood processor only through upstream
feed and so fails the direct test unless palm is itself a processing
input, while the underlying deforestation-commodity exposure is in any
case already binding through EUDR at Tier 1 (soy certifications such as
RTRS and ProTerra are excluded on the same reasoning). Fourth, the
investor-reporting frameworks are consolidated: SASB and TCFD are not
listed separately but are incorporated into IFRS S1/S2 under the ISSB,
and CSRD and ESRS are held as a single entry. General business
regulation that is neither food integrity nor sustainability - the VAT
Directive and the Union Customs Code - is excluded.

The map is deliberately not mutually exclusive. Because Phase 3
decomposes each requirement into its underlying disclosure topics in
order to locate overlap, a topic demanded by several instruments is left
visible under each of them rather than assigned to a single owner.

Each instrument is documented with its issuing organisation, year
established or effective, EU regulatory status, sector applicability,
geographic scope, primary stakeholder driver, obligation locus,
seafood-sector relevance, references to cohort report pages, and
official URL. This produces the landscape-mapping reference document
that accompanies this thesis as a structured workbook.

For the detailed Phase 3 demand mapping, a priority subset of
approximately twelve frameworks is selected on the basis of (i) coverage
across the four stakeholder types, (ii) explicit indicator-level demand
on the entity rather than methodological reference, and (iii) presence
in the cohort's actual reporting practice. The provisional priority list
comprises VSME, MSC, ASC, BAP, Sedex/SMETA, IFS Food, BRCGS, EcoVadis,
CDP, SBTi, GRI 13 and GDST.

\subsection{4.5 Phase 3 - Cross-framework demand
mapping}\label{phase-3---cross-framework-demand-mapping}

The central analytical construct is a coverage matrix with rows = 43
disclosure topics and columns = \textasciitilde12 priority frameworks.
Each cell is coded on a three-level scale:

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 0\tabcolsep) * \real{1.0000}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{Coverage scale (defined ex ante)}

0: Not covered: the framework does not address this topic.

1: Touched: the framework mentions or implicitly references the topic,
but does not require specific reporting or certification on it

2: Required: the framework explicitly requires the entity to report on,
certify, or disclose the topic.
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\end{longtable}
}

Every non-zero cell is documented with a one-line justification
including a citation to the source framework paragraph or indicator.
This per-cell documentation constitutes the audit trail and protects
against reviewer challenges to the coding decisions.

The distinction between methodology providers (e.g., GHG Protocol, ISO
14001, which provide accounting or process standards but do not pressure
entities to report) and disclosure askers (e.g., CDP, SBTi, MSC -which
require entities to submit information) is made explicit.
Methodology-only frameworks are documented in the landscape mapping but
treated separately in the demand analysis.

\subsection{4.6 Phase 4 - Hotspot analysis and stage
assignment}\label{phase-4---hotspot-analysis-and-stage-assignment}

Once the demand matrix is populated, several analytical operations are
performed.

\textbf{Aggregation by stakeholder type}

The framework columns are collapsed into four stakeholder-driver buckets
(regulator-SME, regulator-sector, customer, investor, industry/NGO),
producing for each disclosure topic a count of frameworks demanding it
within each stakeholder type.

\textbf{Hotspot scoring}

Two complementary measures are computed per topic:

\begin{itemize}
\item
  \textbf{Coverage breadth:} total number of priority frameworks with a
  non-zero coverage code for the topic.
\item
  \textbf{Stakeholder diversity:} number of distinct stakeholder-driver
  types in which at least one framework requires the topic. This measure
  rewards cross-stakeholder convergence over within-stakeholder
  repetition.
\end{itemize}

Hotspot identification prioritises stakeholder diversity, on the
conceptual argument that topics demanded by multiple stakeholder types
deliver the highest leverage from a single capability investment.

\textbf{Stage assignment}

Topics are assigned to one of three development stages, with thresholds
set after inspection of the empirical distribution:

\begin{itemize}
\item
  \textbf{Stage 1 - Foundational:} topics with $\geq$3 stakeholder types
  demanding and $\geq$6 frameworks covering. Multiple stakeholders
  independently identify these as material; reporting infrastructure
  built for one stakeholder serves several. Highest priority for SME
  initial investment.
\item
  \textbf{Stage 2 - Expansion:} topics with 2 stakeholder types and 3--5
  frameworks. Cross-stakeholder consensus is forming; build capability
  as resources allow.
\item
  \textbf{Stage 3 - Specialist / regulatory-only:} topics covered by
  ESRS regulation or by one specialist framework only. Required for
  compliance where in scope, but not yet drawing broad demand.
\end{itemize}

\textbf{The diagnostic tool}

The diagnostic tool operationalises the stage-of-development logic for
individual SMEs. It comprises two inputs and two outputs.

\textbf{Input 1 - current reporting posture self-assessment:} a short
structured self-assessment (5--8 items) positioning the SME on the
reactive-to-proactive spectrum. Indicative items: presence of formal
sustainability strategy; presence of a dedicated sustainability function
or role; whether reporting is initiated externally (reactive) or
internally (proactive); presence of voluntary commitments (e.g.,
SBTi-aligned targets); whether disclosure data is reusable across
multiple stakeholder requests.

\textbf{Input 2 - stakeholder-pressure mapping:} the SME identifies
which stakeholders are actively requesting sustainability information
(e.g., specific major retail customers and which questionnaire formats,
specific certifications required for product placement, any investor
pressure, any regulatory pressure beyond CSRD baseline). This input
reflects the SME\textquotesingle s actual external demand environment.

\textbf{Output 1 - prioritised disclosure topics:} the subset of Stage 1
topics whose demanding frameworks align with the SME\textquotesingle s
identified stakeholder pressures. These are the topics for which
capability investment delivers maximum response coverage given the
SME\textquotesingle s specific situation.

\textbf{Output 2 - staged roadmap:} a recommended sequencing of
capability investments across Stage 1 $\to$ Stage 2 $\to$ Stage 3, scaled to the
SME\textquotesingle s current posture and resourcing.

The diagnostic positioning on the reactive-to-proactive spectrum is
operationalised through the self-assessment in Input 1, on the basis
that reactive vs proactive is a behavioural rather than a measurable
categorical state, and that the SME itself is best placed to
characterise its current orientation.

\subsection{4.7 Phase 5 - Case study
validation}\label{phase-5---case-study-validation}

The framework is validated through one or two deep case studies of
European mid-sized seafood processing SMEs (50--250 employees, primarily
B2B suppliers to large EU retailers). For each case the following
sequence is applied:

\begin{itemize}
\item
  Initial scoping interview (\textasciitilde60 min) to characterise the
  firm: size, operations, value-chain position, existing certification
  portfolio, current sustainability reporting practice, current
  customer/investor pressures.
\item
  Application of the diagnostic tool --- the firm completes the
  self-assessment and stakeholder-pressure mapping under researcher
  guidance.
\item
  Generation of the recommended prioritised topic set and staged roadmap
  from the framework.
\item
  Validation interview (\textasciitilde60 min) to discuss the
  recommendations: feasibility, perceived utility, gaps, misalignments
  with the firm\textquotesingle s lived context. Findings inform
  refinement of the tool.
\end{itemize}

The case study sample is intentionally small (1--2 firms) and the
validation is formative rather than confirmatory. Generalisation beyond
the case is not claimed; the case studies are used to test usability and
refine the diagnostic tool, not to validate the underlying empirical
claims about the cohort or the cross-framework demand pattern.

\subsection{4.8 Framework validation
strategy}\label{framework-validation-strategy}

The seafood sustainability playing field (Section 3.4) is a conceptual
framework, and the standards by which such a framework is assessed
differ from those applied to a hypothesis. A conceptual framework is not
proven; it is developed through a structured process and substantiated
through convergent evidence (Jabareen, 2009; Meredith, 1993). The
framework advances two distinct claims that require different forms of
substantiation. The descriptive claim is that the seafood sustainability
disclosure environment can be exhaustively and unambiguously organised
by the necessity, domain and source axes. The explanatory claim, which
is the more substantive and the more demanding to substantiate, is that
capability corridors exist, such that capability investment compelled by
a lower-tier mandatory requirement materially subsidises higher-tier
strategic disclosure. Five complementary validation strategies are
employed; no single one is treated as conclusive, and the framework's
credibility rests on their convergence.

\subsubsection{Leg 1 - Theoretical
grounding}\label{leg-1---theoretical-grounding}

Each axis of the framework is anchored in established theory rather than
derived ad hoc. The necessity axis draws on legitimacy theory (Suchman,
1995) and institutional theory (DiMaggio \& Powell, 1983); the source
overlay on stakeholder-salience theory (Mitchell, Agle \& Wood, 1997);
the capability-corridor construct on the resource-based view (Barney,
1991) and the dynamic-capabilities literature (Teece, 2007); and the
domain axis on materiality and sector-standard literature. Demonstrating
that every structural component of the framework follows from prior
theory substantiates its architecture as theoretically coherent.

\subsubsection{Leg 2 - Descriptive completeness and inter-coder
reliability}\label{leg-2---descriptive-completeness-and-inter-coder-reliability}

The descriptive claim is tested empirically during Phase 2. Every
instrument identified in the landscape mapping is classified into
exactly one necessity-by-domain cell with one primary source
attribution. The framework is descriptively valid to the extent that
this classification is exhaustive, in that no instrument requires a
residual category, and unambiguous, in that instruments genuinely
spanning cells are few and are documented as such. To establish that the
classification is replicable rather than idiosyncratic, a second coder
independently classifies a sample of instruments, and inter-coder
agreement is reported using Cohen's kappa; agreement above the
conventional threshold for substantial reliability substantiates that
the framework can be applied consistently.

\subsubsection{Leg 3 - Expert face
validity}\label{leg-3---expert-face-validity}

The framework is presented to expert informants, namely sustainability
leads at the cohort companies and managers at the SME case-study firms,
who are asked whether its tiers, domains and corridors correspond to
their professional experience. This constitutes member checking, a
recognised qualitative-validation technique (Lincoln \& Guba, 1985).
Expert feedback is documented systematically, including instances in
which it challenges the framework, and revisions made in response to
such feedback are reported as evidence of validation rather than
concealed.

\subsubsection{Leg 4 - Substantiation of the capability-corridor
mechanism}\label{leg-4---substantiation-of-the-capability-corridor-mechanism}

The explanatory claim is substantiated through two converging methods.
First, an input-overlap analysis codes the underlying data and process
inputs required by each instrument situated on a corridor; demonstrating
that, for example, MSC chain-of-custody, the GDST standard and ESRS
value-chain disclosure draw on a common body of lot-level traceability
data provides objective evidence that the corridor is real. Second,
retrospective testimony is gathered from the large-reporter cohort:
informants are asked directly whether capability built for a mandatory
requirement reduced the marginal effort of subsequent disclosures.
Convergence between the objective input-overlap evidence and the
testimonial evidence substantiates the corridor mechanism more strongly
than either method alone.

\subsubsection{Leg 5 - Instrumental validation through case
application}\label{leg-5---instrumental-validation-through-case-application}

Finally, the framework is applied to the SME case studies described in
Section 4.7. If it correctly locates a firm's current posture, correctly
identifies the corridors offering that firm the greatest leverage given
its stakeholder pressures, and is recognised by the firm as an accurate
and actionable diagnosis, the framework is validated instrumentally,
that is, by its demonstrated usefulness in application. This corresponds
to the pattern-matching logic of case-study research (Yin, 2018): the
framework predicts a pattern, and the case evidence is examined for
correspondence to it.

Together, these five strategies validate the framework to a degree
appropriate to the available evidence, and it is important to state the
limits of this claim precisely. The descriptive claim, that the
framework organises the observed ecosystem, can be substantiated
robustly through Legs 1 to 3. The explanatory claim, that mandatory
compliance subsidises strategic sustainability through capability
corridors, is substantiated as plausible and evidenced through Leg 4 and
as useful through Leg 5 but, given the small cohort and the limited
number of case studies, is not claimed to be causally proven. The
defensible position is that the evidence is consistent with the corridor
mechanism, that firms report having experienced it, and that the
instruments on a corridor demonstrably share inputs, rather than that a
causal relationship has been established beyond doubt.

\section{5. Expected outputs and chapter
structure}\label{expected-outputs-and-chapter-structure}

The thesis produces five concrete artefacts mapping to chapter structure
as follows.

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3205}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3397}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3397}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{Chapter}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Central artefact}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Form}
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Ch. 1 Introduction + Ch. 2 Conceptual framework & The reactive$\to$proactive
frame; the cross-stakeholder convergence argument & Prose; reactive vs
proactive comparison table \\
Ch. 3 Methodology & The five-phase methodology, the 43-topic universe
definition, the 0/1/2 coverage scale & Prose; methodology figure showing
the funnel from 70 DRs $\to$ 36 $\to$ 43 \\
Ch. 4 Cohort empirical analysis & Descriptive empirics of large-reporter
disclosure practice & Heatmaps; per-topic and per-company coverage
charts \\
Ch. 5 Cross-framework demand mapping & The central heatmap (33 topics $\times$
37 frameworks); stakeholder aggregation & Master heatmap figure;
per-stakeholder breakdown charts \\
Ch. 6 Stages-of-development roadmap + diagnostic tool & Stage 1/2/3
assignment per topic; the diagnostic tool architecture & Staged roadmap
figure; diagnostic tool process diagram \\
Ch. 7 Case study validation & Application and refinement based on 1--2
SME case studies & Case study narratives with tool output \\
Ch. 8 Discussion + policy implications & Implications for SME
sustainability practice; engagement with the EU simplification debate
(Omnibus, VSME, sector standards) & Prose; brief policy
recommendations \\
Appendix & Prioritised KPI shortlist for European seafood SMEs (Stage 1
+ Stage 2 disclosures expanded to representative KPIs) & Structured
table \\
Appendix & Full landscape mapping (37 frameworks with metadata) &
Reference table \\
Appendix & Full coverage matrix (43 $\times$ 12 with per-cell justifications) &
Audit table \\
\end{longtable}
}

\section{}\label{section}

\section{}\label{section-1}

\section{6. Scope notes and known
limitations}\label{scope-notes-and-known-limitations}

\begin{itemize}
\item
  \textbf{Empirical scope.} Findings apply to European mid-sized seafood
  processing SMEs (50--250 employees) primarily supplying large EU
  retailers under B2B arrangements. Extension to other SME profiles
  (smaller firms, primary producers, B2C operations) is treated as
  future work.
\item
  \textbf{Cohort representativeness.} The five-company cohort represents
  an upper bound of sector disclosure depth rather than an average.
  Inferences about typical practice are not made from this cohort; the
  cohort is used to establish what is empirically feasible to disclose,
  not what is typical.
\item
  \textbf{Cross-framework coding subjectivity.} Coding of framework
  coverage on the 0/1/2 scale involves analyst judgement. Mitigation:
  ex-ante coding rubric, per-cell documentation with source citations,
  sensitivity analysis on threshold choices.
\item
  \textbf{Case study generalisability.} One or two case studies cannot
  validate the framework empirically across the SME population. The case
  studies serve as usability tests of the diagnostic tool, not as
  confirmatory empirical evidence for the framework structure.
\item
  \textbf{Static framework snapshot.} The regulatory and voluntary
  framework landscape is evolving rapidly (Omnibus simplification, VSME
  refinement, CSDDD phased entry, sector ESRS postponement). Findings
  reflect the landscape as observed in May 2026.
\end{itemize}

APPENDIX:

TOOL

\subsubsection{What the SME sees - Step 1: Self-assessment (Input
1)}\label{what-the-sme-sees---step-1-self-assessment-input-1}

\textbf{Where are you on the reactive $\leftrightarrow$ proactive spectrum? (7 items,
takes 5 minutes)}

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.0198}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.6397}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3405}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
\textbf{\#}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\textbf{Question}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\textbf{SME\textquotesingle s answer}
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1 & Do you have a formal written sustainability strategy approved by
leadership? & In development \\
2 & Do you have a dedicated sustainability function or full-time role? &
Part of another role (50\%) \\
3 & What primarily initiates your sustainability reporting work? &
Mostly external requests, some internal \\
4 & Do you have voluntary external commitments (e.g., SBTi, Net-Zero
pledge)? & Considering --- not yet committed \\
5 & Is your sustainability data reusable across multiple customer
requests? & Partially --- rebuilt for each major customer \\
6 & Do you measure and report on metrics that no customer or regulator
currently requires? & No \\
7 & Does sustainability factor into your 3-year strategic planning? &
Limited --- mentioned but not central \\
\end{longtable}
}

$\to$ \textbf{Score: 9 / 21 --- early-stage reactive, transitioning.}

\subsubsection{Step 2: Active stakeholder pressures (Input
2)}\label{step-2-active-stakeholder-pressures-input-2}

\textbf{Tick all that currently apply. Add detail where prompted.}

\emph{Customer pressures (B2B retailer / buyer requirements):}

\begin{itemize}
\item
  $\checkmark$ Sedex/SMETA audit required $\to$ \emph{by Lidl, ICA}
\item
  $\checkmark$ EcoVadis assessment requested $\to$ \emph{by Albert Heijn}
\item
  $\checkmark$ MSC chain-of-custody $\to$ \emph{all wild-capture products}
\item
  $\checkmark$ ASC certification (in progress) $\to$ \emph{for farmed salmon launch}
\item
  $\checkmark$ Customer ESG questionnaire (custom format) $\to$ \emph{Lidl annual; ICA
  biannual}
\item[$\square$]
  BRCGS / IFS Food (food safety) --- \emph{already certified, taken as
  given}
\end{itemize}

\emph{Investor / financial pressures:}

\begin{itemize}
\item
  $\checkmark$ Bank sustainability-linked loan covenant $\to$ \emph{small green loan
  from local bank, requires basic emissions reporting}
\item[$\square$]
  CDP submission requested
\item[$\square$]
  Equity investor ESG questions
\end{itemize}

\emph{Regulatory pressures (direct):}

\begin{itemize}
\item[$\square$]
  CSRD direct (out of scope)
\item
  $\checkmark$ VSME considering for next reporting year
\item
  $\checkmark$ EU IUU regulation compliance (catch certificates for sourced wild
  fish)
\item
  $\checkmark$ EU Control Regulation 2023/2842 --- digital traceability rollout
\end{itemize}

\emph{Industry / NGO / Sector:}

\begin{itemize}
\item[$\square$]
  GRI reporting
\item[$\square$]
  Industry coalition / FAIRR engagement
\item[$\square$]
  NGO scrutiny
\end{itemize}

\subsubsection{Step 3: What the tool does (the
logic)}\label{step-3-what-the-tool-does-the-logic}

The tool now cross-references:

\begin{itemize}
\item
  The SME\textquotesingle s \emph{active frameworks} (from Input 2):
  Sedex/SMETA, EcoVadis, MSC, ASC, Lidl-questionnaire,
  ICA-questionnaire, bank-loan-covenant, VSME, IUU, Control Reg
\item
  The demand matrix (43 topics $\times$ 12 frameworks): which topics each of
  those active frameworks asks about
\item
  Stage assignments per topic
\end{itemize}

Filters and ranks the Stage 1 topics by \emph{how many of the
SME\textquotesingle s active frameworks demand each topic}.

\subsubsection{Step 4: What the SME gets back ---
Output}\label{step-4-what-the-sme-gets-back-output}

\textbf{Your situation:} \emph{Early-stage reactive (9/21). Active
demand from 6 customer-driven frameworks, 1 investor-driven, 4
regulatory.}

\textbf{Output 1 --- Your top 10 priority topics (Stage 1, ranked by
demand alignment with your pressures):}

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.0629}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.3051}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.4353}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.1967}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
\textbf{Rank}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\textbf{Topic}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\textbf{\# of your active frameworks demanding this}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\textbf{Stakeholder types covered}
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1 & Scope 1 + 2 GHG emissions & 7 of your 11 (Sedex, EcoVadis, MSC
chain-of-custody, Lidl Q, VSME, bank-loan, \ldots) & Customer, Investor,
Regulator \\
2 & Own workforce H\&S incidents & 6 (Sedex, EcoVadis, Lidl Q, ICA Q,
ASC, VSME) & Customer, Regulator \\
3 & Chain-of-custody / product traceability & 6 (MSC, ASC, Lidl Q, IUU,
Control Reg, ICA Q) & Customer, Regulator \\
4 & Energy consumption & 5 (EcoVadis, Lidl Q, VSME, bank-loan, ASC) &
Customer, Investor, Regulator \\
5 & Water use & 4 (ASC, EcoVadis, Lidl Q, VSME) & Customer, Regulator \\
6 & Value chain worker policies (forced labour, child labour) & 4
(Sedex/SMETA, ICA Q, EcoVadis, Lidl Q) & Customer \\
7 & Anti-corruption policy + incidents & 4 (Sedex, EcoVadis, Lidl Q,
VSME) & Customer, Regulator \\
8 & Waste generated by processing & 4 (Lidl Q, EcoVadis, VSME, ASC) &
Customer, Regulator \\
9 & Animal welfare (aquaculture-specific) & 3 (ASC, BAP, NGO interest) &
Customer \\
10 & Antimicrobial use (aquaculture) & 3 (ASC, BAP, F2F Strategy) &
Customer, Regulator \\
\end{longtable}
}

\textbf{Output 2 --- Your staged roadmap (personalised):}

\textbf{Stage 1 --- Foundation (start now, 6--12 months):} Build
capability on topics 1--5. These satisfy $\geq$4 of your active frameworks
each. Recommended sequence: GHG (1) $\to$ traceability (3) $\to$ energy + water
(4, 5) $\to$ H\&S (2). Estimated coverage gain: \textasciitilde70\% of your
current external requests answerable from one data infrastructure.

\textbf{Stage 2 --- Expansion (12--24 months, once Stage 1 is
operational):} Add topics 6--8 plus VSME basic module formalisation.
This positions you to handle Lidl\textquotesingle s deepening
sustainability scorecard and EcoVadis level improvements, plus respond
to additional customer ESG questionnaires without rebuild.

\textbf{Stage 3 --- Differentiation (24+ months, proactive moves):}
Topics 9--10 (animal welfare, antimicrobial use) --- currently demanded
by ASC + sector frameworks but not yet by most retailers. Investing here
positions Nordic Whitefish as an aquaculture sustainability leader as
you launch farmed salmon, ahead of likely retailer asks in 2027--28.

\textbf{Positioning note:}

Your self-assessment placed you at 9/21 --- \emph{early-stage reactive
transitioning}. The Stage 1 priorities above, if implemented, would move
you to \emph{capability-driven proactive}, because your reporting
infrastructure would be designed to serve multiple stakeholder demands
simultaneously rather than rebuilt per request.
\appendix
\section{Ecosystem Glossary}
\input{annex.tex}

\end{document}
