% ===== References — generated from BIB.rtf (Zotero export). =====
% RUBEN: 10 inline cites still missing from your library — Adams, Barney,
%   EFRAG, ISO, Jabareen, Lincoln, Philips, Russo, Teece, Yin. Add in Zotero,
%   re-export, and I will regenerate this file.
\phantomsection
\addcontentsline{toc}{section}{References}
\section*{References}
\begingroup
\setlength{\parindent}{0pt}\setlength{\parskip}{6pt}\small
\everypar{\hangindent=1.5em\hangafter=1}

Ahmad, Khursheed, Maria Grazia Marchesano, Valentina Popolo, Roberto Revetria, and Anastasiia Rozhok. ``Development of a European Sustainability Reporting Standards Compliant Sustainability Assessment Framework for Manufacturing Organisations Using Analytic Hierarchy Process.'' \emph{Sustainability} 17, no. 11 (2025): 4772. https://doi.org/10.3390/su17114772.

Ashurst. ``European Commission Consults on Revised ESRS and Sustainability Reporting Standard for Voluntary Use.'' Accessed May 13, 2026. https://www.ashurst.com/en/insights/european-commission-consults-on-revised-esrs-and-sustainability-reporting-standard-for-voluntary-use/.

Barclay, Kate, and Alice Miller. ``The Sustainable Seafood Movement Is a Governance Concert, with the Audience Playing a Key Role.'' \emph{Sustainability} 10, no. 1 (2018): 180. https://doi.org/10.3390/su10010180.

Bataleblu, Ali Asghar, Erwin Rauch, David S. Cochran, and Dominik T. Matt. ``Impact of European Sustainability Reporting Standards Guidelines on the Design of Sustainable Factories and Manufacturing Systems.'' In \emph{Advances in Manufacturing IV}, edited by Adam Hamrol, Marta Grabowska, and Marcin Hinz. Lecture Notes in Mechanical Engineering. Springer Nature Switzerland, 2024. https://doi.org/10.1007/978-3-031-56474-1\_18.

Battisti, Martina, and Martin Perry. ``Walking the Talk? Environmental Responsibility from the Perspective of Small-Business Owners.'' \emph{Corporate Social Responsibility and Environmental Management} 18, no. 3 (2011): 172--85. https://doi.org/10.1002/csr.266.

Baumgartner, Rupert J., and Daniela Ebner. ``Corporate Sustainability Strategies: Sustainability Profiles and Maturity Levels.'' \emph{Sustainable Development} 18, no. 2 (2010): 76--89. https://doi.org/10.1002/sd.447.

Bush, Simon, and Peter Oosterveer. \emph{Governing Sustainable Seafood}. Routledge, 2019. https://doi.org/10.4324/9781315780429.

Chan, Ricky Y. K., Jennifer W. M. Lai, and Namwoon Kim. ``Strategic Motives and Performance Implications of Proactive versus Reactive Environmental Strategies in Corporate Sustainable Development.'' \emph{Business Strategy and the Environment} 31, no. 5 (2022): 2127--42. https://doi.org/10.1002/bse.3011.

Cöster, Mathias, Gunnar Dahlin, and Raine Isaksson. ``Are They Reporting the Right Thing and Are They Doing It Right?---A Measurement Maturity Grid for Evaluation of Sustainability Reports.'' \emph{Sustainability} 12, no. 24 (2020): 10393. https://doi.org/10.3390/su122410393.

DiMaggio, Paul J., and Walter W. Powell. ``The Iron Cage Revisited: Institutional Isomorphism and Collective Rationality in Organizational Fields.'' \emph{American Sociological Review} 48, no. 2 (1983): 147--60. https://doi.org/10.2307/2095101.

Dinh, Tami, Anna Husmann, and Gaia Melloni. ``Corporate Sustainability Reporting in Europe: A Scoping Review.'' \emph{Accounting in Europe} 20, no. 1 (2023): 1--29. https://doi.org/10.1080/17449480.2022.2149345.

Fazli, Haleem, Sami Farooq, Cheng Yang, and Brian Vejrum Wæhrens. ``Proactive and Reactive Approaches towards Sustainable Practices in Manufacturing Companies: Emerging Economies Perspective.'' \emph{Sustainability} 15, no. 17 (2023): 12796. https://doi.org/10.3390/su151712796.

``Feedback on Sustainability Reporting Standards: Additional Explanatory Information Regarding the Value Chain Cap - Finance.'' Accessed May 13, 2026. https://finance.ec.europa.eu/news/feedback-sustainability-reporting-standards-additional-explanatory-information-regarding-value-chain-2026-05-06\_en.

Fragidis, Garyfallos, and Triantafyllos Papafloratos. ``An Information Architecture for the European Sustainability Reporting Standards (ESRS).'' \emph{Sustainability} 17, no. 17 (2025): 7675. https://doi.org/10.3390/su17177675.

Himanen, Satu. \emph{Master's Programme in Circular Economy, Master´s Thesis}. 2025.

Hummel, Katrin, and Dominik Jobst. ``An Overview of Corporate Sustainability Reporting Legislation in the European Union.'' \emph{Accounting in Europe} 21, no. 3 (2024): 320--55. https://doi.org/10.1080/17449480.2024.2312145.

Lozano, Rodrigo. ``Are Companies Planning Their Organisational Changes for Corporate Sustainability? An Analysis of Three Case Studies on Resistance to Change and Their Strategies to Overcome It.'' \emph{Corporate Social Responsibility and Environmental Management} 20, no. 5 (2013): 275--95. https://doi.org/10.1002/csr.1290.

Nahrstedt, Jan. \emph{Double Materiality in Practice: Evidence from Two Years of CSRD Reporting}. n.d.

Nielsen, Christian. ``ESG Reporting and Metrics: From Double Materiality to Key Performance Indicators.'' \emph{Sustainability} 15, no. 24 (2023): 16844. https://doi.org/10.3390/su152416844.

Sharma, Sanjay, and Harrie Vredenburg. ``Proactive Corporate Environmental Strategy and the Development of Competitively Valuable Organizational Capabilities.'' \emph{Strategic Management Journal} 19, no. 8 (1998): 729--53.

Suchman, Mark C. ``Managing Legitimacy: Strategic and Institutional Approaches.'' \emph{The Academy of Management Review} 20, no. 3 (1995): 571--610. https://doi.org/10.2307/258788.

Torugsa, Nuttaneeya Ann, Wayne O'Donohue, and Rob Hecker. ``Proactive CSR: An Empirical Analysis of the Role of Its Economic, Social and Environmental Dimensions on the Association between Capabilities and Performance.'' \emph{Journal of Business Ethics} 115, no. 2 (2013): 383--402. https://doi.org/10.1007/s10551-012-1405-4.

Tugliani, Teresa, Marianna Guareschi, and Filippo Arfini. \emph{Evaluating the Applicability of the European Voluntary Sustainability Reporting Standards for Agrifood Small and Medium-Sized Enterprises}. n.d.

World Wildlife Fund. ``What Are Fishery Improvement Projects?'' Accessed June 4, 2026. https://www.worldwildlife.org/our-work/oceans/improving-fisheries/what-are-fishery-improvement-projects-and-how-do-they-work/.

\endgroup
